---
title: "Lesson 1: Prospective Memory Decision Control"
author: "Dr. Luke Strickland"
date: "`r Sys.Date()`"
output:
  pdf_document: default
  html_document: default
---

The primary requirement for these scripts is the EMC2 package, 
which can be installed from CRAN with install.packages("EMC2").Tidyverse packages 
are also used for utility functions - specifically dplyr, tidyr, ggplot2, and 
stringr. These packages can all be installed from CRAN with install.packages.

## Overview

This script demonstrates how to specify a cognitive process EAM within linear-model language in EMC2.
Specifically, the Prospective Memory Decision Control (PMDC) model includes psychological processes 
that map to linear combinations of parameters, embedded in contrasts. 

We demonstrate how to fit PMDC application to a lexical-decision task example, in which participants
would be expected to decide whether letter strings are words or non-words.  The PM task
could be, for example, to identify any word that is an animal. 

This script defines contrasts to convert psychological processes from PMDC into
"mapped" parameters for the LBA likelihood. The contrast equations are represented by a matrix, 
which will later be supplied to a design object for the R package EMC2. The script also goes through simulation
and evaluation of model parameters, to demonstrate a model validation workflow.


#### Key Psychological Processes in the PMDC Model

The key PMDC processes determining accumulation rates include:

  1) **Ongoing-task quality (qual):**
  - The difference between the matching ("correct") ongoing task accumulation rate and
  mismatching accumulation rate.
  - *Indexes how good ongoing-task processing is (discrimination quality)* 

  2) **Ongoing-task urgency (urg):**
  - The sum of matching and mismatching ongoing-task accumulation rates.
  - *Indexes how fast processing is.*

  3) **Feedforward inhibition induced by PM inputs (inh):** 
  - The difference between ongoing-task accumulation rates in non-PM vs. PM trials.
  - *Indexes the extent to which PM inputs slow the accumulation of the OT.* 
  
  4) **Proactive Control**
  - Adaptations to thresholds to meet PM demands. 
  - For example, an increase in ongoing-task thresholds in PM conditions relative to control.
  - *Measures how participants strategically delay ongoing-task responding to make more time for PM accumulation/retrieval*


Load packages, as well as some bespoke utility functions.
```{r results='hide', message=FALSE}
#Primary package for model fitting
library(EMC2)
#Data manipulation
library(dplyr)
#Some bespoke utility functions
source("utility_functions.R")
load("img/rate_contrasts.RData")
load("data/simple_data.RData")
```


We use data from Strickland et al. (2018, Psychological Review) to get an EMC2 data structure representative
of a PM experiment:

*subjects*, an EMC standard for subject id\
*cond* for blocked condition (control/PM), a factor specific to this experiment\
*S* for stimulus type (w - word, n - non-word p - PM[which is a type of word])\
*R* for response (W - word, N - nonword, P - PM)\
*rt*, RT in seconds

*subjects*, *S*, *R* and *rt* are standard expected EMC2 variable names

```{r}
head(dats)
```


## Contrasts

For this model, we use a custom contrast matrix to embed the accumulation-rate based processes from PMDC.
This matrix is defined in *Lesson1_contrasts.R*. In this contrast matrix, 

  - **Rows** define linear combinations required for each "mapped" accumulation rate.
  - **Columns** define the parameters used to create the mapped accumulation rates.
  
We  create 4 quality/urgency pairs (one for each condition x stimulus type), 
defining the column and row names for the matrix. We also include three PM-specific parameters, 
the PM accumulation rate, PM-induced inhibition of ongoing-task accumulation,
and the PM "false alarm" accumulation rate on non-PM trials.\

#### **PM Accumulator**
We theorise about the PM accumulator rate directly. So, the equation 
for the PM accumulator rate is mapped directly from the PM trial rate.
A single false alarm rate is estimated for both PM stimuli types *(Not of much interest due to poor estimation. Typical studies observe very few PM fas)*.


#### **Inhibition**
To model inhibition, we theorise about the difference between ongoing task-accumulation rates on PM trials, compared to nonPM trials. Inhibition from PM inputs should lead to a lower ongoing task accumulation rate on PM trials.\

The inhibition effect is calculated as the difference between ongoing-task accumulation rates in PM trials compared to non-PM trials. 

  - Ongoing task rate [PM trial] = Ongoing task rate [non-PM trial] - Inhibition

Thus, in the contrast matrix, inhibition is subtracted from the ongoing-task accumulation rates on PM trials:
To model inhibition from PM inputs, we subtract inhibition from ongoing-task accumulation on PM trials:


#### **Ongoing task capacity**
Ongoing task capacity inferred from accumulation rates are decomposed into:
   
   - Urgency (matching + mismatching)
   - Quality (matching - mismatching)\

So:
  
  - Matching rate = (0.5 x Urgency) + (0.5 x Quality) 
  - Mismatching rate = (0.5 x Urgency)  - (0.5 x Quality) 

This structure allows us to estimate capacity separately for PM and control conditions. In this example, PM items are words and so they get the same mappings of ongoing-task parameters as words do.


The contrast matrix and corresponding equations can be viewed below.
```{r}


rate_design

for(i in rownames(rate_design)){
  print(read_rows(rate_design, i))
}
```


## Simulate

Using the contrast specifications, and a data frame
defining a plausible data structure (*data/simple_data.R*), the below code simulates
synthetic performance data from PMDC.


#### **Design Function** 
The design function is a key workhorse of EMC2. For documentation see `?design()`. Here we use it to specify a PM model with: 
  
  - 19 parameters
  - 11 parameters determining the mean accumulation rates
  - 5 parameters determining threshold
  - 1 Start-point noise parameter
  - 1 Non-decision time parameter
  - 1 fixed sv parameter (sv incorrect responses), and one determining sv for correct responses
  
Proactive control is embedded in the specification of thresholds:`B ~ cond*lR`. This
linear-model syntax reflects that for B we will estimate a threshold intercept, main effects of PM condition
and latent response and the interaction between them. Overall proactive control over the ongoing task would
be reflected by a main effect of condition. In this study, PM targets were always words. Targeted
proactive control would be represented by an interaction between lR and cond, with higher thresholds specifically
for the competing word accumulator in PM blocks. 

B is by default estimated on the log scale in EMC2, to avoid negative threshold values where less evidence
is required to respond than at the start point. This means that the parameter is exponentiated before
applied to the model likelihood. Thus, B should be interpreted in terms of multiplicative effects: values
greater than one imply a relative increase in thresholds, and less than one imply a decrease. sv, t0, and A
are all also estimated on the log scale to enforce positivity.
  

```{r results='hide'}


#This match function identifies correct responses (*used here only for sv*):
match_fun <- function(d)
  as.character(d$S) == tolower(
    as.character(d$lR)
  )


design_PM_sim <- design(
  # LBA has historically been used to specify PMDC theory, but other race models
  #could also be used such as the Racing Diffusion Model.
  model=LBA, 
  #The dummy data frame we will use for simulation
  data=dats,
  # Functions can create new experimental factors, which are used here to map 
  # out contrast equations. We create an SlR factor that matches the rownames from
  #the defined contrast matrix (in 1-specify_contrasts.R).
  functions=list(SlR=function(d) 
    factor(paste0(d$lR, d$S, d$cond),
  #It is important the order is the same so we need to give it these levels
           levels=rownames(rate_design))
    ,
  # The RACE argument is used to deal with the fact that in PM designs, 
  #control conditions have 2 accumulators, and PM conditions have three. 
  #The below input assures that the PM accumulator is excluded from control 
  #blocks. 
    RACE=function(d)factor(ifelse(d$cond=="C",2,3)
    )
  ),
  #Here the contrast matrix is supplied
  contrasts = list(SlR=rate_design),
  # A match function is always expected, even when the factor it creates (lM) 
  #is not used by design (so it can be used in functions that summarize accuracy).
  matchfun=match_fun,
  # This lm-style formula describes how parameters are mapped. Key here: **v** 
  # is mapped according to SlR, which in turn uses the supplied contrast matrix. 
  # To illustrate a conventional mapping, **B** here is just mapped using 
  #linear-model notation. Specifically, it    #can vary by condition and by 
  #latent accumulator (lR). Increased ongoing-task thresholds in PM, as compared 
  #with control, conditions corresponds to "proactive control" in PMDC theory. 
  #These effects will be given by the "condPM" and "condPM:lRW" effects in this
  # example
  formula=list(v~SlR,B ~ cond*lR,A ~ 1,t0 ~ 1,sv ~ lM),
  #Enforce a bound for inhibition >0
  pre_transform = list(func = c(v_SlRinh = "exp")),
  constants=c(
    # sv set to 1 here as a scaling parameter
    # To force it to be positive, EMC estimates log(sv) then exponentiates 
    # afterwards. Hence set to log(1)
    sv=log(1),
    # By default, an intercept parameter is also included, which we don't want 
    # as the contrasts already fully define the v parameters for this example. 
    #So set the intercept to constant 0
    v=0,
    # There is no lRP in the PM condition (no third accumulator). 
    # So, it would redundant to have both lRP and condPMxlRP included in 
    # model estimation.     # Set condPM x lRP fixed at log(1) to avoid this issue.
    # Inputting as log(1) rather than 0 is just a reminder that thresholds are 
    # estimated on the log scale (to avoid negative results)
    "B_condPM:lRP" = log(1)
  )
)

```
Verify mapping equations:
```{r}
mapped_pars(design_PM_sim)
```

#### **Run Simulation**

Simulation values were determined by fitting the model to data
from Strickland et al. (2018), Experiment 1. Specifically, we fit control
and non-focal PM conditions, excluding focal, to get estimates of PMDC parameters.

From this we extracted population-mean parameter estimates, and a covariance matrix indexing parameter variance across participants and relationships between parameters (see Stricklandetal_2018_specify.R and Stricklandetal_check_savevalues.R).

Here, the EMC2 function (`make_random_effects`) uses the covariance matrix to create 
random effects for parameters, simulating parameters for each p that can then in turn be used to simulate from PMDC. The result is a 35 (subject) x 19 (row) matrix.

The `make_data` function uses the participant parameters to simulate data frame matching
that supplied.

```{r}

load("data/sim_values.RData")

res_recovery <- 
  make_random_effects(
    design= design_PM_sim, 
    group_means = p_vector_sim,
    covariances = hierarchical_covs
  )

res_recovery[,"t0"] <- pmax(res_recovery[,"t0"], log(0.05))
```

```{r eval=FALSE}
simmed_group <- make_data(res_recovery, data=dats, design = design_PM_sim)

# save(
#   simmed_group, design_PM_sim, res_recovery, p_vector_sim,
#   hierarchical_covs,
#   file="data/simmed_group.RData"
# )

```


## Make Samplers

EMC2 uses "sampler" objects to fit models, which includes: 

  - **Data**: The observed or simulated dataset.
  - **Design**: The experiment structure and parameter mappings.
  - **Priors**: Initial distributions for parameter estimation.

This section specifies samplers to fit the synthetic data that we just generated. Specifically:
  - One sampler for a model matching the structure of the "true" generating model.
  - Another sampler incorrectly excluding proactive-control effects.

Below recreates the design using the simulated data. The only difference is that we use different data (simulated rather than dummy) to extract information about experiment factors./
```{r results='hide'}
load("data/simmed_group.RData")

design_PM <- design(model=LBA, 
                    #The data created last script
                    data=simmed_group,
                    functions=list(SlR=function(d) 
                      factor(paste0(d$lR, d$S, d$cond),
                             levels=rownames(rate_design)),
                      RACE=function(d)factor(ifelse(d$cond=="C",2,3)
                      )
                    ),
                    pre_transform = list(func = c(v_SlRinh = "exp")),
                    contrasts = list(SlR=rate_design),
                    matchfun=match_fun,
                    formula=list(v~SlR,B ~ cond*lR,A ~ 1,t0 ~ 1,sv ~ lM),
                    constants=c(sv=log(1),v=0,
                                "B_condPM:lRP" = log(1)
                    )
)
```

#### **Define Priors**
Bayesian estimation requires priors. To specify priors, EMC2 requires values for population means and standard
deviations for each parameter. A multivariate normal distribution is assumed across participants,
with sensible default priors for correlations.

Priors are plotted below. Also, mapped population-mean values are printed to ensure correct mappings
via the mapped_pars function.
  
```{r}
p_vector <- sampled_pars(design_PM,doMap=FALSE)

p_vector[grepl("A", names(p_vector))] <- log(0.2)
p_vector[grepl("B", names(p_vector))] <- log(1)
p_vector["t0"] <- log(0.2)

p_vector[grepl("inh", names(p_vector))] <- log(0.1)
p_vector[grepl("v_SlRPMR", names(p_vector))] <- 2

p_vector[grepl("qual", names(p_vector))] <- 2
p_vector[grepl("urg", names(p_vector))] <- 2



n_vs <- length(p_vector[grepl("^v", names(p_vector))])
n_Bs <- length(p_vector[grepl("B", names(p_vector))])

psd=c(
  #rates
  rep(2,n_vs),
  #thresholds,
  rep(1,(n_Bs)),
  #A, t0, sv
  1, 1, 1
  
)

prior_PM  <- prior(design_PM,mu_mean=p_vector, mu_sd=psd)
plot(prior_PM,map=F,layout=c(2,5))

mapped_pars(design_PM, p_vector)
```


A sampler emc object must be created for model fitting. Save the sampler object for later use. 
We dispatched the fit at the terminal from *run_simulate_fit.R*.
```{r eval=FALSE}

emc <- make_emc(simmed_group, design_PM, prior=prior_PM)
# save(emc,file="samples/samplerPM.RData")
```

#### **Alternative Model Without Proactive-Control Effects**
This model excludes the condition effect on **B**, testing an alternative explanation (hypothesis) against the primary model.\

The design specifications are similar.
```{r results='hide'}
design_PM_noproactive <- design(model=LBA, 
                                data=simmed_group,
                                functions=list(SlR=function(d) 
                                  factor(paste0(d$lR, d$S, d$cond),
                                         levels=rownames(rate_design)),
                                  RACE=function(d)factor(ifelse(d$cond=="C",2,3)
                                  )
                                ),
                                pre_transform = list(func = c(v_SlRinh = "exp")),
                                contrasts = list(SlR=rate_design),
                                matchfun=match_fun,
                                # Note here B ~ lR rather than B ~ cond*lR
                                formula=list(v~SlR,B ~ lR,A ~ 1,t0 ~ 1,sv ~ lM),
                                constants=c(sv=log(1),v=0
                                )
)

prior_PM_noproactive  <- prior(design_PM_noproactive,update=prior_PM)

```

```{r eval=FALSE}

emc <- make_emc(simmed_group, design_PM_noproactive, prior=prior_PM_noproactive)
# save(emc,file="samples/samplerPM_noproactive.RData")
```


## Examine Simulation
Below we examine parameter recovery for the fitted model. That is, how well the parameter estimates obtained via estimation
match the estimates used to generate the data. This tests how well the model "worked" in estimating the correct values given the data structure, in an example where the "truth" is known.

#### **Load Data and Model Samples**
Load the simulated dataset, rate contrast matrix, and finished model samples. Look at Rhat/ESS (converged and enough effective samples?)

Generally, everything is well converged (Rhat<1.1). Effective sample size is good (>1000) for the mu parameters,
with the exception of the PM false alarm rate (which is poorly estimated) and the threshold-intercept (but not threshold effects). More samples for these parameters could be collected by running `fit` for longer, but we do not do so here.

```{r}
load("img/rate_contrasts.RData")
load("data/simmed_group.RData")
fitPM <- get(load("samples/samplerPM.RData"))
fitPMnc <- get(load("samples/samplerPM_noproactive.RData"))
mus <- summary(fitPM, selection="mu")
```

Ensure that parameter estimation converged successfully:
```{r eval=FALSE}
check(fitPM)
check(fitPMnc)

```


#### **Parameter Recovery** 
Here we check the recovery of population-level parameters by comparing estimates to the original simulated values.Both the population mean (mu) and individual level estimates (alpha) look reasonable, as indicated by a good fit
to the y=x line.

```{r}
print(round(recovery(fitPM,layout=NULL,
               p_vector_sim,
               selection="mu"
)$mu$quantiles,2))


```

```{r}

recovery(fitPM,
               res_recovery,
               selection="alpha"
)
```


#### **Model Recovery**
Another test is whether we can verify that the 'true' generating model structure is supported over alternative explanations, using model comparison tools. For an example, we compare the full model (which includes proactive-control effects) against an alternative model that excludes a proactive-control threshold effect used in the simulation.

We compare models using the Deviance Information Criterion (DIC) and Bayesian Predictive Information Criterion (BPIC). Lower values and higher weights indicate more relative support for a model.

```{r}

compare(list(
  top = fitPM,
  noproactive = fitPMnc
  ), BayesFactor = F
)

```
 
 In this case, the full model outperforms the alternative, suggesting that proactive-control effects are important. Further tests could examine other the effect of excluding other mechanisms such as reactive control, as demonstrated in Lesson2.Rmd. 

