---
title: 'Lesson 2: A unified model of PM and automation use'
author: "Dr. Luke Strickland"
date: "`r Sys.Date()`"
output:
  pdf_document: default
  html_document: default
editor_options:
  markdown:
    wrap: 72
---

The primary requirement for these scripts is the EMC2 package, 
which can be installed from CRAN with install.packages("EMC2").Tidyverse packages 
are also used for utility functions - specifically dplyr, tidyr, ggplot2, and 
stringr. These packages can all be installed from CRAN with install.packages.

## Overview

In this script, we demonstrate how to unify the PMDC model with a recently
published model of automation use. We also step through the basic model testing
and validation workflow. The original work and data evaluated is from
an empirical study:

-   Boag, Strickland, Heathcote & Loft (2025) The Use of Automated Decision Aids in Multi-Tasking Context.*Under Review*.
    
In the proposed model, automation-use affects accumulation rates in two potential ways:

1) Excitation, where automation inputs increase accumulation towards
recommended responses. 
- *Excitation implies that accumulation rates
should be faster than manual rates for the choice automation agrees
with.*
2) Inhibition, where automation inputs decrease accumulation
towards opposing responses. 
- *Inhibition implies accumulation rates should be slower than manual for the choice automation disagrees with.*

As with the PMDC model, the possibility of proactive control over thresholds is included
(larger thresholds in PM than control blocks).
Furthermore, this proactive control is allowed to differ across conditions where participants
perform the task manually versus when they work with automated advice.

Boag et al. (2025) tested the model with an experiment. Participants performed
an air traffic control conflict detection task, where they determined whether 
aircraft pairs would violate safe separation standards during their flight. In some
conditions, participants had a PM task. This required pressing an alternative key
for targets with particular aircraft callsigns. In some conditions, participants were
provided with an automated decision aid that recommends whether
aircraft are in conflict or not. This decision aid was sometimes wrong. PM and 
automation conditions were crossed within-subjects for each participant.

**Boag et al. (2025) Design Factors**

-   *Stimulus Type*: Conflict, Nonconflict, or PM
-   *PM Block*: Control or PM
-   *Automation Block*: Manual (Unaided) or Automation (Aided)
-   *Automation Accuracy*: Correct or Incorrect



## Load

Load packages, bespoke utility functions and relevant data.

```{r results='hide', message=FALSE}
#Primary package for model fitting
library(EMC2)
#Data manipulation
library(dplyr)
#For plots
library(ggplot2)
#Some bespoke utility functions
source("utility_functions.R")

load("data/data_advanced.RData")
load("img/exinhcm_PM.RData")
load("img/Proactivecm_auto.RData")
```



The data set that will be analysed. It contains standard EMC2 columns, such as subjects,S for stimulus type
(cc=conflict, nn=nonconflict, pc= PM target, conflict, pn= PM target, nonconflict),
R for response (C= conflict, N= non-conflict, P=PM), and rt.

Also, design-specific columns:

-   `auto`: Did participants have an automated decision aid in that
    block(auto) or not (manual)?
-   `failtrial`: Did the automated advice recommend a correct decision
    (AC) or incorrect (AI) on the particular trial?
-   `PM`: PM or control conditions?

```{r}
head(dat)
```


## Contrasts

Generally, accumulation-rate contrasts are similar to PMDC except that automation-informed contrasts are added in automation blocks:
-  automation excitation is added to rates agreeing with automated advice (e.g., the matching accumulation rate when
automation is correct)
-  automation inhibition is subtracted from rates disagreeing with automated advice (e.g., the matching 
accumulation rate when automation is incorrect)

These contrasts are additional to the PM inhibition that is subtracted from ongoing task accumulation rates on PM target trials.

 Unlike the last example,
ongoing-task accumulation rates are not specified in terms of "quality"
and "urgency" here, for simplicity. However, they could have been specified that way. 

```{r}
print_full_v_cm <- full_v_cm[apply(full_v_cm, 1, function(row) any(row != 0)), ]
for(i in rownames(print_full_v_cm )){
  print(read_rows(print_full_v_cm , i))
}
```

In this example, proactive control over thresholds will also be coded
using contrasts, rather than included as linear-model effects. It would also 
have been possible to use linear-model syntax to interact automation conditions, PM conditions,
and latent accumulator (lR).

```{r}

print_Proactivecm_auto <- Proactivecm_auto[
  apply(Proactivecm_auto, 1, function(row) any(row != 0)), 
  ]

for(i in rownames(print_Proactivecm_auto)){
  print(read_rows(print_Proactivecm_auto, i))
}
```

## Model Specification


We now demonstrate how to specify a model for this data set. It is similar to the
first example, but requires including extra factors. 

```{r results='hide'}

match_fun <- function(d)
  as.character(substr(d$S,1,1)) == tolower(
    as.character(d$lR)
  )

design_advanced <- design(model=LBA, 
                          # The applied data set
                          data=dat,
                          # The relevant factors for contrasts will match the 
                          # row names in the contrast matrices.
                          functions=list(
                            SlR=function(d) factor(paste0(d$lR, d$S, d$AM, d$fail, d$PM),
                                                   levels=rownames(full_v_cm)),
                            PMlR=function(d) factor(paste0(d$lR, d$PM, d$AM),
                                                    levels=rownames(Proactivecm_auto)),
                            # Special "RACE" factor which will be used to deal with
                            # the fact that PM has different number of accumulators to control
                            RACE=function(d)factor(ifelse(d$PM=="control",2,3))
                          ),
                          # Supplying the contrast matrices
                          contrasts = list(SlR=full_v_cm, PMlR =Proactivecm_auto),
                          matchfun=match_fun,
                          # Specifying how parameters vary, in this case by contrast matrices
                          formula=list(v~SlR,B ~ PMlR,A ~ 1,t0 ~ 1,sv ~ lM),
                          # Appropriate constants
                          constants=c(sv=log(1),v=0, B=log(1))
)
```

Setup a p_vector containing naive prior means

```{r}
p_vector <- sampled_pars(design_advanced)
p_vector[grepl("B", names(p_vector))] <- log(1)
p_vector[grepl("inh", names(p_vector))] <- 0
p_vector[grepl("PMV", names(p_vector))] <- 1
p_vector[grepl("McC", names(p_vector))] <- 1
p_vector[grepl("MnN", names(p_vector))] <- 1
p_vector[grepl("t0", names(p_vector))] <- log(0.4)

n_vs <- length(p_vector[grepl("^v", names(p_vector))])
n_Bs <- length(p_vector[grepl("B", names(p_vector))])

psd=c(
  #drifts
  rep(2,n_vs),
  #Bs,
  rep(1,(n_Bs)),
  #A, t0, sv true
  1, 1, 1
)


```

Create prior and then save the sampler object for later use

```{r eval=FALSE}
prior_advanced  <- prior(design_advanced,mu_mean=p_vector, mu_sd=psd)
emc <- make_emc(dat, design_advanced, prior=prior_advanced)
#save(emc,file="samples/sampler_advanced.RData")
```


#### **Sub-models**

We fit a range of candidate models that might
account for the advanced data set. 

-   "Top" flexible model with full flexibility in PM and automation
    mechanisms specified in the last contrasts scripts. Specified above

-   A model that excludes the PM-inhibition mechanism

-   A model that excludes the PM-induced proactive control mechanisms

-   A model that excludes the automation use mechanisms

These sub-model constraints are implemented by setting parameters to
a constant, but could equally be created by re-specifying contrasts


##### **No PM inhibition**

No PM-inhibition mechanism for either auto or manual conditions. Set the PM inhibition parameters to 0.

```{r results='hide'}
design_advanced_noPMinh <- design(model=LBA, 
                                  data=dat,
                                  functions=list(
                                    SlR=function(d) factor(paste0(d$lR, d$S, d$AM, d$fail, d$PM),
                                                           levels=rownames(full_v_cm)),
                                    PMlR=function(d) factor(paste0(d$lR, d$PM, d$AM),
                                                            levels=rownames(Proactivecm_auto)),
                                    RACE=function(d)factor(ifelse(d$PM=="control",2,3))
                                  ),
                                  contrasts = list(SlR=full_v_cm, PMlR =Proactivecm_auto),
                                  matchfun=match_fun,
                                  formula=list(v~SlR,B ~ PMlR,A ~ 1,t0 ~ 1,sv ~ lM),
                                  constants=c(sv=log(1),v=0, B=log(1),
                                              v_SlRMANPMinh=0,
                                              v_SlRAUTOPMinh=0
                                  )
)

p_vector_noPMinh <-
  p_vector[!names(p_vector) %in% c("v_SlRMANPMinh", "v_SlRAUTOPMinh")]
n_vs_noPMinh <- length(p_vector_noPMinh[grepl("^v", names(p_vector_noPMinh))])
n_Bs_noPMinh <- length(p_vector_noPMinh[grepl("B", names(p_vector_noPMinh))])

psd_noPMinh =c(
  #drifts
  rep(2,n_vs_noPMinh),
  #Bs,
  rep(1,(n_Bs_noPMinh)),
  #A, t0, sv true
  1, 1, 1
)
```

```{r eval=FALSE}
prior_advanced_noPMinh  <- prior(design_advanced_noPMinh,
                                 mu_mean=p_vector_noPMinh, 
                                 mu_sd=psd_noPMinh
)
emc <- make_emc(dat, design_advanced_noPMinh, prior=prior_advanced_noPMinh)
#save(emc,file="samples/sampler_advanced_noPMinh.RData")
```


##### **No Proactive Control**

No proactive control. Set the proactive control parameters to log(1) (i.e., 0).

```{r results='hide'}
design_advanced_noproactive <- design(model=LBA, 
                                      data=dat,
                                      functions=list(
                                        SlR=function(d) factor(paste0(d$lR, d$S, d$AM, d$fail, d$PM),
                                                               levels=rownames(full_v_cm)),
                                        PMlR=function(d) factor(paste0(d$lR, d$PM, d$AM),
                                                                levels=rownames(Proactivecm_auto)),
                                        RACE=function(d)factor(ifelse(d$PM=="control",2,3))
                                      ),
                                      contrasts = list(SlR=full_v_cm, PMlR =Proactivecm_auto),
                                      matchfun=match_fun,
                                      formula=list(v~SlR,B ~ PMlR,A ~ 1,t0 ~ 1,sv ~ lM),
                                      constants=c(sv=log(1),v=0, B=log(1), 
                                                  B_PMlRmanualpaC=log(1),
                                                  B_PMlRmanualpaN=log(1),
                                                  B_PMlRautopaC=log(1),
                                                  B_PMlRautopaN=log(1)
                                      )
)


p_vector_noproactive <-
  p_vector[!names(p_vector) %in% c("B_PMlRmanualpaC", "B_PMlRmanualpaN",
                                   "B_PMlRautopaC", "B_PMlRautopaN")]
n_vs_noproactive <- length(p_vector_noproactive[grepl("^v", names(p_vector_noproactive))])
n_Bs_noproactive <- length(p_vector_noproactive[grepl("B", names(p_vector_noproactive))])


psd_noproactive =c(
  #drifts
  rep(2,n_vs_noproactive),
  #Bs,
  rep(1,(n_Bs_noproactive)),
  #A, t0, sv true
  1, 1, 1
)
```

```{r eval=FALSE}
prior_advanced_noproactive  <- prior(design_advanced_noproactive,
                                     mu_mean=p_vector_noproactive, 
                                     mu_sd=psd_noproactive
)
emc <- make_emc(dat, design_advanced_noproactive, prior=prior_advanced_noproactive)

#save(emc,file="samples/sampler_advanced_noproactive.RData")
```

##### **No Automation Use Mechanisms**

Exclude automation-use mechanisms by fixing such parameters to 0.

```{r results='hide'}
design_advanced_noauto <- design(model=LBA, 
                                 data=dat,
                                 functions=list(
                                   SlR=function(d) factor(paste0(d$lR, d$S, d$AM, d$fail, d$PM),
                                                          levels=rownames(full_v_cm)),
                                   PMlR=function(d) factor(paste0(d$lR, d$PM, d$AM),
                                                           levels=rownames(Proactivecm_auto)),
                                   RACE=function(d)factor(ifelse(d$PM=="control",2,3))
                                 ),
                                 contrasts = list(SlR=full_v_cm, PMlR =Proactivecm_auto),
                                 matchfun=match_fun,
                                 formula=list(v~SlR,B ~ PMlR,A ~ 1,t0 ~ 1,sv ~ lM),
                                 constants=c(sv=log(1),v=0, B=log(1),
                                             v_SlRPMAUTOex = 0, v_SlRPMAUTOinh =0,
                                             v_SlRcontrolAUTOex = 0, v_SlRcontrolAUTOinh =0
                                 )
)

p_vector_noauto <-
  p_vector[!names(p_vector) %in% c("v_SlRPMAUTOex", "v_SlRPMAUTOinh",
                                    "v_SlRcontrolAUTOex", "v_SlRcontrolAUTOinh")]

n_vs_noauto <- length(p_vector_noauto[grepl("^v", names(p_vector_noauto))])
n_Bs_noauto <- length(p_vector_noauto[grepl("B", names(p_vector_noauto))])

psd_noauto =c(
  #drifts
  rep(2,n_vs_noauto),
  #Bs,
  rep(1,(n_Bs_noauto)),
  #A, t0, sv true
  1, 1, 1
)
```


```{r eval=FALSE}
prior_advanced_noauto  <- prior(design_advanced_noauto,
                                mu_mean=p_vector_noauto, 
                                mu_sd=psd_noauto
)
emc <- make_emc(dat, design_advanced_noauto, prior=prior_advanced_noauto)

#save(emc,file="samples/sampler_advanced_noauto.RData")
```

## Model Evaluation

Model workflow often involves a break for fitting times. Reload necessary
functions. Tidyverse functions for summarizing data, predictions and plotting.

```{r}
library(tidyr)
library(dplyr)
library(ggplot2)
library(EMC2)
theme_set(theme_bw())
source("utility_functions.R")
load("img/exinhcm_PM.RData")
load("img/Proactivecm_auto.RData")
```

Load some utility functions not packaged in EMC but helpful for
summarizing output

-   Specifically, functions that take data and posterior predictives and
    create summaries for fit plots
-   EMC2 also has some built-in options, but the options here make plots
    most consistent with previous publications of the PMDC model


Load the top model and reload the necessary contrast matrices. Check the 
top model converged adequately. 

```{r}
fit_advanced <- get(load("samples/sampler_advanced.RData"))
```


#### **Evaluate Posterior Predictive Fit**

Get posterior predictives using EMC2's predict method. Save (can be slow and better
to dispatch multi-core).
```{r eval=FALSE}
check(fit_advanced)
pp <- predict(fit_advanced, n_cores=1)
dat <- get_data(fit_advanced)
#save(pp, dat, file= "samples/predicted_advanced.RData")
```


```{r}
load("samples/predicted_advanced.RData")
```

EMC2 provides summary and plotting functions. However, is it also straightforward
to extract data and posterior predictives to specify custom plots, each are provided
simply as data frames. Here, we aim to 
produce plots consistent with Boag et al. (2025). To that end, a non-EMC2 utility function,
`calc_stat_postfit` takes `dat` and `pp` extracted from the EMC2 fit and summarizes
based on a provided function.

The custom function requires a data summary specified as function (here, accuracy),
experimental factors to examine, and `stat_cols` that are required for the function (here S and R 
are required to calculate accuracy). It returns summaries of the data and posterior predictives.
By default, the grouped data frame is examined, providing information about average fit. The
"subjects" column can also be applied to look at individual participants.

```{r}
acc_fun <- function(S, R){
  mean(substr(S,1,1)==tolower(R))
}

acc_df <- calc_stat_postfit(dat, pp, facs=c("S", "AM", "fail", "PM"),
                            stat_fn=acc_fun,
                            stat_cols=c("S","R")
)

acc_df$AM_fail <- factor(
  paste0(acc_df$AM, acc_df$fail),
  levels=c( "manualAC", "autoAC","autoAI"),
  labels=c("manual", "AC", "AI"))

```



Plots below visualize accuracy across
automation and failure conditions (`AM_fail`), split by stimulus type
(`S`) and PM condition.

-   The first plot focuses on `"cc"` and `"nn"` trials, faceted by `S`
    and `PM`.
-   The second plot excludes those trials and shows remaining `S` types
    side-by-side
-   Dashed lines connect grouped conditions (`nam`)
-   The black point layer (`stat`) provides comparison values (e.g.,
    model predictions or alternative metrics)

As plotted below, fits to both PM and ongoing-task accuracy look quite good. Fits are generally improved in 
the Boag et al. (2025) paper, which specified a more complex model.

```{r}
acc_df %>% filter(S %in% c("cc","nn")) %>% ggplot(aes(AM_fail, postmn_stat)) +
  geom_point(size=3)+ geom_errorbar(aes(ymax = upper, ymin = lower), width=0.5) +
  geom_line(aes(group=1), linetype=2)+
  geom_point(aes(y= stat), pch=21, size=4, colour="black")+
  facet_grid(S~PM)  +ylab("Correct RT") +xlab("Automation") +ylab("Accuracy")+
  theme(text = element_text(size = 22), axis.text=element_text(size=15))+ylim(0.5,1)
```

```{r}
acc_df %>% filter(!S %in% c("cc","nn")) %>% ggplot(aes(AM_fail, postmn_stat)) +
  geom_point(size=3)+ geom_errorbar(aes(ymax = upper, ymin = lower), width=0.5) +
  geom_line(aes(group=1), linetype=2)+
  geom_point(aes(y= stat), pch=21, size=4, colour="black")+
  facet_grid(.~S)  +ylab("Correct RT") +xlab("Automation") +ylab("Accuracy")+
  theme(text = element_text(size = 22), axis.text=element_text(size=15)) +ylim(0.5,1)
```

For RT, we wish to examine three aspects of RT distribution: 0.1 quantile,
median (0.5 quantile), and 0.9 quantile. This gives an indiciation of distributional 
fit. Another non-EMC2 convenience function `stack_stats_postfits` can apply 
to get multiple summaries of data simultaneously and bind to a data frame.

Requires:

-   A list of functions (`corRT_fn_list`, defined earlier) to compute
    the quantiles
    `corRT_fn_list` computes the 10th,
50th, and 90th percentiles of reaction times for error trials
-   The columns to pass each function (`cols`). Equivalent to stat_cols in the earlier function
-   Names of the resulting statistics for the outputted data frame

```{r}
corRT_fn_list <- c(function(S, R, rt){quantile(rt[toupper(substr(S,1,1))==R], probs=0.1, na.rm=T)},
                   function(S, R, rt){quantile(rt[toupper(substr(S,1,1))==R], probs=0.5, na.rm=T)},
                   function(S, R, rt){quantile(rt[toupper(substr(S,1,1))==R], probs=0.9, na.rm=T)}
)

cols <- c("S", "R", "rt")
stat_col_list <- list(cols,cols,cols)
stat_fn_names <- c("0.1", "0.5", "0.9")

cRT_df <- stack_stat_postfits(
  dat, pp, facs=c("S", "AM", "fail", "PM"), corRT_fn_list, stat_col_list, stat_fn_names
) 

cRT_df$AM_fail <- factor(
  paste0(cRT_df$AM, cRT_df$fail),
  levels=c( "manualAC", "autoAC","autoAI"),
  labels=c("manual", "AC", "AI"))
```

Plots below show good distributional fits to correct RT, other than a small miss
on the 0.9 quantile for the rare PM response times.

```{r}
cRT_df %>% filter(S %in% c("cc","nn")) %>%  ggplot(aes(AM_fail, postmn_stat)) +
  geom_point(size=3)+ geom_errorbar(aes(ymax = upper, ymin = lower), width=0.5) +
  geom_line(aes(group=nam), linetype=2)+
  geom_point(aes(y= stat), pch=21, size=4, colour="black")+
  facet_grid(S~PM)  +ylab("Correct RT")  +xlab("Automation")+
  theme(text = element_text(size = 22), axis.text=element_text(size=15))

cRT_df %>% filter(!S %in% c("cc","nn")) %>%  ggplot(aes(AM_fail, postmn_stat)) +
  geom_point(size=3)+ geom_errorbar(aes(ymax = upper, ymin = lower), width=0.5) +
  geom_line(aes(group=nam), linetype=2)+
  geom_point(aes(y= stat), pch=21, size=4, colour="black")+
  facet_grid(.~S)  +ylab("Correct RT")  +xlab("Automation")+
  theme(text = element_text(size = 22), axis.text=element_text(size=15))
```


Error RTs visualized similarly below, except for PM trials due to this being a 
rare type of response. 

```{r}
errRT_fn_list <- c(function(S, R, rt){quantile(rt[toupper(substr(S,1,1))!=R],
                                               probs=0.1)},
                   function(S, R, rt){quantile(rt[toupper(substr(S,1,1))!=R], 
                                               probs=0.5)},
                   function(S, R, rt){quantile(rt[toupper(substr(S,1,1))!=R], 
                                               probs=0.9)}
)

eRT_df <- stack_stat_postfits(
  dat, pp, facs=c("S", "AM", "fail", "PM"), errRT_fn_list, stat_col_list, stat_fn_names
)

eRT_df$AM_fail <- factor(
  paste0(eRT_df$AM, eRT_df$fail),
  levels=c( "manualAC", "autoAC","autoAI"),
  labels=c("manual", "AC", "AI"))


eRT_df %>% filter(S %in% c("cc","nn")) %>% ggplot(aes(AM_fail, postmn_stat)) + 
  geom_point(size=3)+ geom_errorbar(aes(ymax = upper, ymin = lower), width=0.5) +
  geom_line(aes(group=nam), linetype=2)+
  geom_point(aes(y= stat), pch=21, size=4, colour="black")+
  facet_grid(S~PM)  +ylab("Error RT") +
  theme(text = element_text(size = 22), axis.text=element_text(size=15))
```

#### **Model Comparison**

Test whether the model structure can be simplified. Smaller DIC/BPIC suggest
a better fit to that model. DIC/BPIC also help to determine the relative contributions of different
mechanisms to overall fit - larger differences in favour of a model suggest a stronger preference.


```{r}
fit_advanced_noauto <- 
  get(load("samples/sampler_advanced_noauto.RData"))
fit_advanced_noPMinh <- 
  get(load("samples/sampler_advanced_noPMinh.RData"))
fit_advanced_noproactive <- 
  get(load("samples/sampler_advanced_noproactive.RData"))
```

All submodels converged in terms of Rhat <1.1 all parameters.
```{r eval=FALSE}
check(fit_advanced_noauto)
check(fit_advanced_noPMinh)
check(fit_advanced_noproactive)
```


Below uses EMC2's compare function to examine whether dropping mechanisms is supported.
Both DIC and BPIC suggest that the top model, with full PM and
automation-use mechanisms, is most supported. Thus, all the
included mechanisms are useful for predicting the data.

Bayes Factors could also have been calculated *by setting BayesFactor=TRUE), but they would require 
more posterior samples and were not used by Boag et al. (2025). S,o we have not done so here.

```{r}
out <- compare(list(
  top = fit_advanced,
  noPMinh = fit_advanced_noPMinh,
  noproactive = fit_advanced_noproactive,
  noauto = fit_advanced_noauto
), BayesFactor = FALSE)

out$DIC <- out$DIC - out$DIC[1]
out$BPIC <- out$BPIC - out$BPIC[1]
out
```

#### **Model Parameter Summaries**

Summarise the model parameters to understand the underlying mechanisms implied.

1.  Extract population mean parameters from the full model
2.  Plot the posterior means/ 95% credible intervals
3.  Examine Bayesian p-values and Savage-Dickey Bayes Factors to evaluate effects

EMC2's get_pars function extracts parameters, here we focus on 'mu' (population means).
```{r}
params <- data.frame(
  get_pars(fit_advanced, selection="mu", merge_chains = T)$mu[[1]]
)
```

First, examine the excitation and inhibition mechanisms, both involving
automation and the PM task. 

This reveals:
-  strong PM inhibition in both auto and manual conditions
-  strong automation inhibition in the PM condition, but less in control
-  excitation strategy strong in control conditions, but not PM condition

```{r}
param_summaries <- params %>% pivot_longer(cols=colnames(params)) %>% 
  group_by(name) %>% 
  summarize(mn=mean(value), 
            p025= quantile(value, probs=0.025),
            p975=quantile(value, probs=0.975)
  )

exinh <- param_summaries  %>% filter(grepl(".*[ex|inh]$", name))

exinh$name <- c("PM inhibition, auto condition", 
                "PM inhibition, manual condition",
                "Automation excitation, PM condition",
                "Automation inhibition, PM condition",
                "Automation excitation, control condition",
                "Automation inhibition, control condition"
)

exinh  %>% 
  ggplot(aes(x=name, y=mn, ymin=p025, ymax=p975))+
  ylab("Effect")+ xlab("") +
  geom_pointrange()+ 
  geom_hline(yintercept=0, lty=2) +  # add a dotted line at x=1 after flip
  coord_flip() +
  theme(text = element_text(size = 18))


```

Calculating Bayesian p-value (posterior sample density <0) and 
Savage-Dickey test (Bayes Factor of alternative vs null of 0) using EMC2's
`credible` and `hypothesis` functions. 
set.seed because there is some noise in sampling from the prior.
```{r results='hide', fig.show = "hide"}

set.seed(1234)
PMInh_auto_p <- attr(
  credible(x = fit_advanced, x_name = "v_SlRAUTOPMinh"), "less"
)
PMInh_auto_SD <- hypothesis(fit_advanced, parameter="v_SlRAUTOPMinh",  do_plot = F)


PMInh_man_p <- attr(
  credible(x = fit_advanced, x_name = "v_SlRMANPMinh"), "less"
)
PMInh_man_SD <- hypothesis(fit_advanced, parameter="v_SlRMANPMinh",  do_plot = F)


AutoInh_PM_p <- attr(
  credible(x = fit_advanced, x_name = "v_SlRPMAUTOinh"), "less"
)
AutoInh_PM_SD <- hypothesis(fit_advanced, parameter="v_SlRPMAUTOinh",  do_plot = F)

AutoInh_control_p <- attr(
  credible(x = fit_advanced, x_name = "v_SlRcontrolAUTOinh"), "less"
)
AutoInh_control_SD <- hypothesis(fit_advanced, parameter="v_SlRcontrolAUTOinh",  do_plot = F)


AutoEx_PM_p <- attr(
  credible(x = fit_advanced, x_name = "v_SlRPMAUTOex"), "less"
)
AutoEx_PM_SD <- hypothesis(fit_advanced, parameter="v_SlRPMAUTOex",  do_plot = F)


AutoEx_control_p <- attr(
  credible(x = fit_advanced, x_name = "v_SlRcontrolAUTOex"), "less"
)
AutoEx_control_SD <- hypothesis(fit_advanced, parameter="v_SlRcontrolAUTOex",  do_plot = F)

#Differences are more complex so use bigger N
PM_control_AutoInh_p <- attr(
  credible(x = fit_advanced, x_fun = function(d) d["v_SlRPMAUTOinh"]
         - d["v_SlRcontrolAUTOinh"]),
  "less"
)
PM_control_AutoInh_SD <- hypothesis(
  fit_advanced, fun=function(d) d["v_SlRPMAUTOinh"]
           - d["v_SlRcontrolAUTOinh"], do_plot = F)

PM_control_AutoEx_p <- attr(
  credible(x = fit_advanced, x_fun = function(d) d["v_SlRcontrolAUTOex"] 
         - d["v_SlRPMAUTOex"]), "less"
)

PM_control_AutoEx_SD <- hypothesis(fit_advanced, fun= function(d) 
  exp(d["v_SlRcontrolAUTOex"]) - exp(d["v_SlRPMAUTOex"]), do_plot = F)
```


```{r}

effnames <- c("PM inhibition, manual", "PM inhibition, auto",
              "Auto inhibition, PM", "Auto inhibition, control",
              "Auto excitation, PM", "Auto excitation, control",
              "Auto inhibition, PM - Auto inhibition, control",
              "Auto excitation, PM - Auto excitation, control")

ps <- c(PMInh_auto_p, PMInh_man_p,AutoInh_PM_p, AutoInh_control_p,
        AutoEx_PM_p, AutoEx_control_p, PM_control_AutoInh_p, 
        PM_control_AutoEx_p)

SDs <- c(PMInh_auto_SD, PMInh_man_SD,AutoInh_PM_SD, AutoInh_control_SD,
        AutoEx_PM_SD, AutoEx_control_SD, PM_control_AutoInh_SD, 
        PM_control_AutoEx_SD)

df <- data.frame(Effect=effnames, p=round(ps, 3), SD= round(SDs, 3))

df

```
Now examine proactive control. 
This reveals:
- strong proactive control of both accumulators in both conditions
- weak tendency to less control in automation conditions

```{r}
proactive <- param_summaries %>% filter(grepl(".*pa.$", name))

proactive$name <- c("Conflict Accumulator, Automation", 
                    "Non-conflict Accumulator, Automation", 
                    "Conflict Accumulator, Manual", 
                    "Non-conflict Accumulator, Manual" 
)

proactive  %>% 
  ggplot(aes(x=name, y=mn, ymin=p025, ymax=p975))+
  ylab("Proactive Control Effect")+ xlab("") +
  geom_pointrange()+ 
  geom_hline(yintercept=0, lty=2) +  # add a dotted line at x=1 after flip
  coord_flip() +theme(text=element_text(size=14)) 

```

Bayesian p values and SD tests.
```{r results="hide", fig.show = "hide"}

set.seed(1234)
ProAutoC_p <- attr(credible(x = fit_advanced, x_name= "B_PMlRautopaC"),
                 "less")
ProAutoC_SD <- hypothesis(fit_advanced, parameter="B_PMlRautopaC",  do_plot = F)


ProAutoN_p <- attr(credible(x = fit_advanced, x_name= "B_PMlRautopaN"),
                 "less")
ProAutoN_SD <- hypothesis(fit_advanced, parameter="B_PMlRautopaN",  do_plot = F)



ProManC_p <- attr(credible(x = fit_advanced, x_name= "B_PMlRmanualpaC"),
                 "less")
ProManC_SD <- hypothesis(fit_advanced, parameter="B_PMlRmanualpaC",  do_plot = F)


ProManN_p <- attr(credible(x = fit_advanced, x_name= "B_PMlRmanualpaN"),
                 "less")
ProManN_SD <- hypothesis(fit_advanced, parameter="B_PMlRmanualpaN",  do_plot = F)



ProManAutoC_p <-  attr(credible(x = fit_advanced, x_fun = function(d) d["B_PMlRmanualpaC"] - d["B_PMlRautopaC"]
         ), "less"
)

ProManAutoC_SD <-  hypothesis(fit_advanced, fun = function(d) d["B_PMlRmanualpaC"] - 
             d["B_PMlRautopaC"],  do_plot = F
)



ProManAutoN_p <-  attr(credible(x = fit_advanced, x_fun = function(d) d["B_PMlRmanualpaN"] - d["B_PMlRautopaN"]
         ), "less"
)

ProManAutoN_SD <-  hypothesis(fit_advanced, fun = function(d) d["B_PMlRmanualpaN"] - 
             d["B_PMlRautopaN"],  do_plot = F
)


```

Simple proactive-control effects are large, but note the interaction (manual - auto)
is quite weak, with only a tendency for increase in manual not supported by p value
or SD test.
```{r}

effnames <- c("Conflict, Automation", 
              "Non-conflict, Automation", 
               "Conflict, Manual", 
              "Non-conflict, Manual",
              "Conflict, Manual - Conflict, Automation",
              "Non-conflict, Manual - Non-conflict, Automation")

ps <- 
  c(ProAutoC_p, ProAutoN_p, ProManC_p, ProManN_p, ProManAutoC_p,ProManAutoN_p)

SDs <- 
  c(ProAutoC_SD, ProAutoN_SD, ProManC_SD, ProManN_SD, ProManAutoC_SD,ProManAutoN_SD)

df <- data.frame(Effect=effnames, p=round(ps, 3), SD= round(SDs, 3))

df

```

New proactive control plot and new p-values (~0.01, ~0.02 for conflict and non-conflict,
respectively).

```{r}
set.seed(1234)

params_mapped <- data.frame(
  get_pars(fit_advanced, selection="mu", merge_chains = T, map=T)$mu[[1]]
)

params_mapped$ProAC <- params_mapped$B_PMlRCPMauto - params_mapped$B_PMlRCcontrolauto 
params_mapped$ProAN <- params_mapped$B_PMlRNPMauto - params_mapped$B_PMlRNcontrolauto 

params_mapped$ProMC <- params_mapped$B_PMlRCPMmanual - params_mapped$B_PMlRCcontrolmanual 
params_mapped$ProMN <- params_mapped$B_PMlRNPMmanual - params_mapped$B_PMlRNcontrolmanual

params_mapped$ProMAC <- params_mapped$ProMC - params_mapped$ProAC 
params_mapped$ProMAN <- params_mapped$ProMN - params_mapped$ProAN 


params_mapped_summaries <- params_mapped %>% pivot_longer(cols=colnames(params_mapped )) %>% 
  group_by(name) %>% 
  summarize(mn=mean(value), 
            p025= quantile(value, probs=0.025),
            p975=quantile(value, probs=0.975)
  )


proactive <- params_mapped_summaries %>% filter(grepl("Pro", name) & !(name %in% c("ProMAC", "ProMAN")))

proactive$name <- c("Conflict Accumulator, Automation", 
                    "Non-conflict Accumulator, Automation", 
                    "Conflict Accumulator, Manual", 
                    "Non-conflict Accumulator, Manual" 
)

# Strong proactive control in both auto and manual conditions. Slightly
# stronger for manual, perhaps due to strong automation-inhibition effects 
# providing support in PM conditions

proactive  %>% 
  ggplot(aes(x=name, y=mn, ymin=p025, ymax=p975))+
  ylab("Proactive Control Effect (Additive)")+ xlab("") +
  geom_pointrange()+ 
  geom_hline(yintercept=0, lty=2) +  # add a dotted line at x=1 after flip
  coord_flip() 


#Get p-values and Bayes Factors for the mapped, transformed values
additive_ProManAutoC_p <-  attr(credible(x = fit_advanced, x_fun = function(d) (
 
  (exp(d["B_PMlRmanualC"] +d["B_PMlRmanualpaC"])- exp(d["B_PMlRmanualC"]))  
  - (exp(d["B_PMlRautoC"] +d["B_PMlRautopaC"]) -exp(d["B_PMlRautoC"]))
  )
), "less"
)

additive_ProManAutoC_SD <-  hypothesis(fit_advanced, fun = function(d) (
 
  (exp(d["B_PMlRmanualC"] +d["B_PMlRmanualpaC"])- exp(d["B_PMlRmanualC"]))  
  - (exp(d["B_PMlRautoC"] +d["B_PMlRautopaC"]) -exp(d["B_PMlRautoC"]))
  ),  do_plot = F
)

additive_ProManAutoC_SD 
  
additive_ProManAutoN_p <-  attr(credible(x = fit_advanced, x_fun = function(d) (
 
  (exp(d["B_PMlRmanualN"] +d["B_PMlRmanualpaN"])- exp(d["B_PMlRmanualN"]))  
  - (exp(d["B_PMlRautoN"] +d["B_PMlRautopaN"]) -exp(d["B_PMlRautoN"]))
  )
), "less"
)

additive_ProManAutoN_SD <-  hypothesis(fit_advanced, fun = function(d) (
 
  (exp(d["B_PMlRmanualN"] +d["B_PMlRmanualpaN"])- exp(d["B_PMlRmanualN"]))  
  - (exp(d["B_PMlRautoN"] +d["B_PMlRautopaN"]) -exp(d["B_PMlRautoN"]))
  ),  do_plot = F
)

additive_ProManAutoN_SD


```
