---
title: 'Lesson 1: Design Specification and Model Recovery'
author: "Dr. Luke Strickland"
date: "`r Sys.Date()`"
output:
  pdf_document: default
  html_document: default
  word_document: default
---

## Setup

### Install required packages globally
A straightforward setup is to setup the required packages on your machine.
The primary requirement for these scripts is the EMC2 package, 
which can be installed from CRAN with install.packages("EMC2").Tidyverse packages 
are also used for utility functions - specifically dplyr, tidyr, ggplot2, and 
stringr. These packages can all be installed from CRAN with install.packages.

###Notes

- On windows, some packages may require that "Rtools" is first installed for both
installation methods.

### renv
We also provide an renv, which you can use to automatically install all packages
used for this R project down to the specific versions, specifically
for working with this project alone. To use this, see the script setup_renv.R


## Overview

This script demonstrates how to specify a cognitive process 
evidence accumulation model (EAM) within linear-model language in EMC2.
Specifically, the Prospective Memory (PM) Decision Control (PMDC) model includes psychological processes 
that map to linear combinations of parameters, embedded in contrasts. More information
can be found in the associated manuscript.

We demonstrate how to fit PMDC to a lexical-decision task (the "ongoing" task), in which participants
decide whether letter strings are words or non-words.  The PM task
could be, for example, to identify any word that is an animal. 

This script defines contrasts to convert psychological processes from PMDC into
"mapped" parameters for the LBA likelihood. The contrast equations are represented by a matrix, 
which will later be supplied to a design object for the R package EMC2. The script also goes through simulation and evaluation of model parameters, to demonstrate a model validation workflow.


#### Key Psychological Processes in the PMDC Model

The key PMDC processes determining accumulation rates include:

  1) **Ongoing-task quality (qual):**
  - The difference between the matching ("correct") ongoing task accumulation rate and
  mismatching accumulation rate.
  - *Indexes how good ongoing-task processing is (discrimination quality)* 

  2) **Ongoing-task urgency (urg):**
  - The sum of matching and mismatching ongoing-task accumulation rates.
  - *Indexes how fast processing is.*

  3) **Feedforward inhibition induced by PM inputs (inh):** 
  - The difference between ongoing-task accumulation rates in non-PM vs. PM trials.
  - *Indexes the extent to which PM inputs slow the accumulation of the OT.* 
  
  4) **Proactive Control**
  - Adaptations to thresholds to meet PM demands. 
  - For example, an increase in ongoing-task thresholds in PM conditions relative to control.
  - *Measures how participants strategically delay ongoing-task responding to make more time for PM accumulation/retrieval*


Load packages, as well as some bespoke utility functions.
```{r results='hide', message=FALSE}
#Primary package for model fitting
library(EMC2)
#Data manipulation package
library(dplyr)
#Some bespoke utility functions
source("utility_functions.R")
#A matrix of mapping contrasts which are used to create design-cell
#parameters out of process parameters, for accumulation rates specifically
load("img/rate_contrasts.RData")
#An example data set taken from Strickland et al. (2018)
#Will be used just for the structure (trial number, factors etc.)
load("data/simple_data.RData")
```


We use data from Strickland et al. (2018) to get an EMC2 data structure representative
of a PM experiment:

*subjects*, an EMC standard for subject id\
*cond* for blocked condition (control/PM), a factor specific to this experiment\
*S* for stimulus type (w - word, n - non-word p - PM[which is a type of word])\
*R* for response (W - word, N - nonword, P - PM)\
*rt*, RT in seconds

*subjects*, *S*, *R* and *rt* are standard expected EMC2 variable names

 - Strickland, L., Loft, S., Remington, R. W., & Heathcote, A. (2018). 
 Racing to remember: A theory of decision control in event-based 
 prospective memory. Psychological review, 125(6), 851-887. 
 https://doi.org/10.1037/rev0000113

```{r}
head(dats)
```


## Contrasts

For this model, we use a custom mapping contrast matrix to embed the accumulation-rate based processes from PMDC.
This matrix is defined in *Lesson1_contrasts.R*. In this contrast matrix, 

  - **Rows** define linear combinations required for each mapped accumulation rate.
  - **Columns** define the parameters used to create the mapped accumulation rates.
  
We  create 4 quality/urgency pairs (one for each condition x stimulus type), 
defining the column and row names for the matrix. We also include three PM-specific parameters, 
the PM accumulation rate, PM-induced inhibition of ongoing-task accumulation,
and the PM "false alarm" accumulation rate on non-PM trials.\

#### **PM Accumulator**
We estimate the PM accumulator rate directly ("dummy coding"). So, the equation 
for the PM accumulator rate is mapped directly from the PM trial rate.
A single false alarm rate is estimated for both PM stimuli types *(Not of much interest due to poor estimation. Typical studies observe very few PM fas)*.


#### **Inhibition**
To model inhibition, we theorise about the difference between ongoing task-accumulation rates on PM trials, compared to nonPM trials. Inhibition from PM inputs should lead to a lower ongoing task accumulation rate on PM trials.\

The inhibition effect is calculated as the difference between ongoing-task accumulation rates in PM trials compared to non-PM trials. 

  - Ongoing task rate [PM trial] = Ongoing task rate [non-PM trial] - Inhibition

Thus, in the mapping contrast matrix, inhibition is subtracted from the ongoing-task accumulation rates on PM trials:
To model inhibition from PM inputs, we subtract inhibition from ongoing-task accumulation on PM trials:


#### **Ongoing task capacity**
Ongoing task capacity inferred from accumulation rates are decomposed into:
   
   - Urgency (matching + mismatching)
   - Quality (matching - mismatching)\

So:
  
  - Matching rate = (0.5 x Urgency) + (0.5 x Quality) 
  - Mismatching rate = (0.5 x Urgency)  - (0.5 x Quality) 

This structure allows us to estimate capacity separately for PM and control conditions. In this example, PM items are words and so they get the same mappings of ongoing-task parameters as words do.


The contrast matrix and corresponding equations can be viewed below.
```{r}

#Print the matrix
rate_design

#Loop through it and print out equations with a utility function.
for(i in rownames(rate_design)){
  print(read_rows(rate_design, i))
}
```


## Simulate

Using the mapping contrast specifications, and a data frame
defining a plausible data structure (*data/simple_data.R*), the below code simulates
synthetic performance data from PMDC.


#### **Design Function** 
The design function is a key workhorse of model specification in EMC2. It is used both in
both simulating and fitting data. It is explained extensively
in the manuscript. For further documentation see `?design()`.

Here we specify a PM model with: 
  
  - 19 parameters
  - 11 parameters determining the mean accumulation rates
  - 5 parameters determining threshold
  - 1 Start-point noise parameter
  - 1 Non-decision time parameter
  - 1 fixed sv parameter (sv incorrect responses), and one determining sv for correct responses
  
As described in the manuscript, 
EMC2 applies linear models to map estimated parameters to experiment design cells.
These linear models will apply to a design data frame, which is built internally by the `make_data`
and `make_emc()` functions described below. The design data frame combines a data structure with information from the design function.

The design data frame is expanded to have a row for each trial '
for each accumulator or "latent response" (*lR*). This is provided either by
"Rlevels", or inferred from a data frame supplied to "data".
An additional column called *lM* is also created, to determine which
accumulator matches the correct decision on each trial type. To do so, the user
supplies a match function ("matchfun") outputting TRUE for matches and FALSE for mismatches. In
the PM design, the number of accumulators varies based on conditions. The number of 
accumulators is given by a special argument called RACE, which requires a function
outputting the number of desired accumulators. 

Further custom factor columns not contained in the original data can be
useful for linear modelling. They are added to the design data frame using the "functions"
argument. Here, we will create a custom factor *SlR* that maps to the rows of the mapping
contrast matrix we supplied earlier. This factor will allow EMC to correctly map 
contrasts over the design. These contrasts will be used to specify the mean accumulation rate
parameters, "v".

EMC2 uses the linear-model syntax from R's regression framework.
Proactive control is embedded in the specification of thresholds:`B ~ cond*lR`. This
linear-model syntax reflects "treatment coding" for B in which we will estimate a threshold intercept, main effects of PM condition
and latent response and the interaction between them. Overall proactive control over the ongoing task would
be reflected by a main effect of condition. In this study, PM targets were always words. Targeted
proactive control would be represented by an interaction between *lR* and *cond*, with higher thresholds specifically
for the competing word accumulator in PM blocks. 

By default, some parameters are transformed after mapping in EMC2 to ensure that
they have allowable values. B is by default estimated on the log scale in EMC2, to avoid negative threshold values where less evidence
is required to respond than at the start point. This means that the parameter is exponentiated before
being applied to the model likelihood. Thus, B should be interpreted in terms of multiplicative effects: values
greater than one imply a relative increase in thresholds, and less than one imply a decrease. sv, t0, and A
are all also estimated on the log scale to enforce positivity.
  

```{r results='hide'}

# A match fun is required by EMC2 for data where there is a correct response.
# This match function identifies correct responses:
# Which is used for various functions such as plots.
# It can also be used as a factor in the linear model, in this example
# it is used to allow sv to vary based on whether the accumulator matches
# the correct response or not
match_fun <- function(d)
  as.character(d$S) == tolower(
    as.character(d$lR)
  )


design_PM_sim <- design(
  # LBA has historically been used to specify PMDC theory, but other race models
  #could also be used such as the Racing Diffusion Model.
  model=LBA, 
  #The dummy data frame we will use for simulation
  #design uses this to figure out experiment factors and response levels
  data=dats,
  # Functions can create new experimental factors, which are used here to map 
  # out contrast equations. We create an SlR factor that matches the rownames from
  #the defined contrast matrix (in 1-specify_contrasts.R).
  functions=list(SlR=function(d) 
    factor(paste0(d$lR, d$S, d$cond),
  #It is important the order is the same so we need to give it these levels
           levels=rownames(rate_design))
    ,
  # The RACE argument is used to deal with the fact that in PM designs, 
  #control conditions have 2 accumulators, and PM conditions have three. 
  #The argument below ensures that the PM accumulator is excluded from control 
  #blocks. 
    RACE=function(d)factor(ifelse(d$cond=="C",2,3)
    )
  ),
  #Here the contrast matrix is supplied
  contrasts = list(SlR=rate_design),
  # It is best to always provide a match function, even when the factor it creates (lM) 
  #is not used by design (so it can be used in functions that summarize accuracy).
  matchfun=match_fun,
  # This lm-style formula describes how parameters are mapped. Key here: **v** 
  # is mapped according to SlR, which in turn uses the supplied contrast matrix. 
  # To illustrate a conventional mapping, **B** here is just mapped using 
  #linear-model notation. Specifically, it can vary by condition and by 
  #latent accumulator (lR). Increased ongoing-task thresholds in PM, as compared 
  #with control, conditions corresponds to "proactive control" in PMDC theory. 
  #These effects will be given by the "condPM" and "condPM:lRW" effects in this
  # example
  formula=list(v~SlR,B ~ cond*lR,A ~ 1,t0 ~ 1,sv ~ lM),
  #Enforce a bound for inhibition >0
  pre_transform = list(func = c(v_SlRinh = "exp")),
  constants=c(
    # sv set to 1 here as a scaling parameter
    # To force it to be positive, EMC estimates log(sv) then exponentiates 
    # afterwards. Hence set to log(1) for clarity (we could also set it to 0)
    sv=log(1),
    # By default, an intercept parameter is also included, which we don't want 
    # as the contrasts already fully define the v parameters for this example. 
    #So set the intercept to constant 0
    v=0,
    # There is no lRP in the PM condition (no third accumulator). 
    # So, it would redundant to have both lRP and condPMxlRP included in 
    # model estimation.     # Set condPM x lRP fixed at log(1) to avoid this issue.
    # Inputting as log(1) rather than 0 is just a reminder that thresholds are 
    # estimated on the log scale (to avoid negative results)
    "B_condPM:lRP" = log(1)
  )
)

```
Verify mapping equations:
```{r}
#Prints out the linear models that will apply across design cells to 
# determine the "mapped" design-cell parameters.
# Note that the intercept v value was fixed at 0, as the accumulation rate
# coding was already fully specified without needing an intercept
mapped_pars(design_PM_sim)
```

#### **Run Simulation**

Having specified design, we will use it to simulate synthetic data from the '
previously described model. 

Simulation requires choosing a set of model parameter
values for each simulated participant. Here, parameter values were chosen by 
fitting the model to real data from Strickland et al. (2018), Experiment 1. 
Specifically, we fit control and non-focal PM conditions, excluding focal, to get estimates of PMDC parameters.  From this we extracted population-mean parameter estimates, and a covariance matrix indexing parameter variance across participants and relationships between parameters (see *Stricklandetal_2018_specify.R* and *Stricklandetal_check_savevalues.R*).We use these estimates here to simulate for illustrative purposes.

We simulate a multi-participant data set
of the same size as Strickland et al. (2018)'s. Here, the EMC2 function 
(`make_random_effects`) uses the aforementioned covariance matrix to create 
random effects for parameters, simulating parameters for each 
p that can then in turn be used to simulate data from PMDC. The result is a 35 (subject) x 19 (row) matrix.

The `make_data` function uses the participant parameters to simulate data frame matching
that supplied. 

```{r}

#Load some parameter values we will simulate from, based on a fit to
# Strickland et al. (2018) Experiment 1
load("data/sim_values.RData")

#Simulate individual-participant parameters using population means and covariances
# obtained from the previous fit
res_recovery <- 
  make_random_effects(
    design= design_PM_sim, 
    group_means = p_vector_sim,
    covariances = hierarchical_covs
  )

#Non-decision time has a bound of 0.05s. Force individual-participant simulated
# parameters above this bound
res_recovery[,"t0"] <- pmax(res_recovery[,"t0"], log(0.05))
```

```{r eval=FALSE}
#This will create simulated choice-RT data using the specified parameters (in this case
# an array of individual-subject parameters), design, and here the data argument
# used to determine data structure, trial numbers etc.
simmed_group <- make_data(parameters= res_recovery, data=dats, design = design_PM_sim)

#Save off all the results of the simulation
# save(
#   simmed_group, design_PM_sim, res_recovery, p_vector_sim,
#   hierarchical_covs,
#   file="data/simmed_group.RData"
# )

```


## Make Samplers

Having simulated synthetic data, we now demonstrate how a model can estimate 
parameters from data. Subsequently we will check whether the estimated values
are a good match to the known "true" simulation values.

EMC2 uses "sampler" objects to fit models, which includes: 

  - **Data**: The observed or simulated dataset.
  - **Design**: The experiment structure and parameter mappings.
  - **Priors**: Initial distributions for parameter estimation.

This section specifies samplers to fit the synthetic data that we just generated. Specifically:
  - One sampler for a model matching the structure of the "true" generating model.
  - Another sampler incorrectly excluding proactive-control effects.

Below recreates the design using the simulated data. 
The only difference from the previous call to design 
is that we use different data (simulated rather than dummy) to extract information about experiment factors./
```{r results='hide'}
load("data/simmed_group.RData")

#See the previous call to design for extensive comments matching these arguments
design_PM <- design(model=LBA, 
                    #The data created last script
                    data=simmed_group,
                    functions=list(SlR=function(d) 
                      factor(paste0(d$lR, d$S, d$cond),
                             levels=rownames(rate_design)),
                      RACE=function(d)factor(ifelse(d$cond=="C",2,3)
                      )
                    ),
                    pre_transform = list(func = c(v_SlRinh = "exp")),
                    contrasts = list(SlR=rate_design),
                    matchfun=match_fun,
                    formula=list(v~SlR,B ~ cond*lR,A ~ 1,t0 ~ 1,sv ~ lM),
                    constants=c(sv=log(1),v=0,
                                "B_condPM:lRP" = log(1)
                    )
)
```

#### **Define Priors**
Bayesian estimation requires priors. To specify priors either for a single participant or for population means (for a hierarchical model, as we focus on here), EMC2 requires values for population means and a standard
deviation representing normally distributed uncertainty. If no prior is provided a standard normal distribution is used. A multivariate normal distribution is assumed across participants,
with sensible default priors for correlations and variances with a limited ability for users to change these settings.

Priors are plotted below. Also, mapped population-mean values are printed to ensure correct mappings
via the mapped_pars function.
  
```{r}
#Note below, for clarity priors on log-scaled parameters are specified
# using log(X)

#A function to get a vector with the right parameter names for the design
p_vector <- sampled_pars(design_PM,doMap=FALSE)

#A, t0, sv match. Reasonable values
p_vector[grepl("A", names(p_vector))] <- log(0.2)
p_vector["t0"] <- log(0.2)
#This implies a prior of match==mismatch
p_vector["sv_lMTRUE"] <- log(1)

#All B values fixed at log (1). This implies a prior on B intercept of 1,
# and a prior on all the B effects of 0
p_vector[grepl("B", names(p_vector))] <- log(1)

#Inhibition prior on a low value. 
p_vector[grepl("inh", names(p_vector))] <- log(0.1)
#PM accumulation rate
p_vector[grepl("v_SlRPMR", names(p_vector))] <- 2

#After mapping these imply a prior of 2 for match rates and 0 for mismatch
# (see call to mapped_pars below)
p_vector[grepl("qual", names(p_vector))] <- 2
p_vector[grepl("urg", names(p_vector))] <- 2

p_vector["v_SlRfa"] <- 0

#Setting up a vector for uncertainty expressed as SDs.

#Number of mean accumulation rates
n_vs <- length(p_vector[grepl("^v", names(p_vector))])
#Number of thresholds
n_Bs <- length(p_vector[grepl("B", names(p_vector))])

psd=c(
  #rates
  rep(2,n_vs),
  #thresholds,
  rep(1,(n_Bs)),
  #A, t0, sv
  1, 1, 1
  
)

#This function creates the mu prior using the design, pop means and uncertainty
# in pop means
prior_PM  <- prior(design_PM,mu_mean=p_vector, mu_sd=psd)
plot(prior_PM,map=F,layout=c(2,5))

#For the prior on pop variance and correlations we use EMC2 defaults, which are 
# sensible

#Below call to mapped_pars (where both design and p vector supplied) will
# show mapped parameters for each design cell
mapped_pars(design_PM, p_vector)
```


A sampler emc object must be created for model fitting. Save the sampler object for later use. 
We dispatched the fit at the terminal from *run_simulate_fit.R*.
```{r eval=FALSE}

#Make emc combines the simulated data and design to make the design data frame.
# Further, it also combines the prior, with the final "emc" object ready to be sampled
emc <- make_emc(simmed_group, design_PM, prior=prior_PM)
#Save off samples and dispatch using script
# save(emc,file="samples/samplerPM.RData")
```

#### **Alternative Model Without Proactive-Control Effects**
This model excludes the condition effect on **B**, testing an alternative explanation (hypothesis) against the primary model.\

The design specifications are similar.
```{r results='hide'}

# A very similar call to design as above, except for a difference in the linear model formula,
# noted below. 
design_PM_noproactive <- design(model=LBA, 
                                data=simmed_group,
                                functions=list(SlR=function(d) 
                                  factor(paste0(d$lR, d$S, d$cond),
                                         levels=rownames(rate_design)),
                                  RACE=function(d)factor(ifelse(d$cond=="C",2,3)
                                  )
                                ),
                                pre_transform = list(func = c(v_SlRinh = "exp")),
                                contrasts = list(SlR=rate_design),
                                matchfun=match_fun,
                                # Note here B ~ lR rather than B ~ cond*lR
                                formula=list(v~SlR,B ~ lR,A ~ 1,t0 ~ 1,sv ~ lM),
                                constants=c(sv=log(1),v=0
                                )
)

# As shown below, the prior function can create a prior for a simpler model when it is nested
# within a more complex model. To do so, supply an argument to update.
prior_PM_noproactive  <- prior(design_PM_noproactive,update=prior_PM)

```

```{r eval=FALSE}

emc <- make_emc(simmed_group, design_PM_noproactive, prior=prior_PM_noproactive)
# save(emc,file="samples/samplerPM_noproactive.RData")
```


## Examine Simulation
Below we examine parameter recovery for the fitted model. That is, how well the parameter estimates obtained via estimation
match the estimates used to generate the data. This tests how well the model "worked" in estimating the correct values given the data structure, in an example where the "truth" is known.

#### **Load Data and Model Samples**
Load the simulated dataset, rate contrast matrix, and finished model samples. Look at Rhat/ESS (converged and enough effective samples?)

Generally, everything is well converged (Rhat<1.1). Effective sample size is good (>1000) for the mu parameters,
with the exception of the PM false alarm rate (which is poorly estimated) and the threshold-intercept (but not threshold effects). More samples for these parameters could be collected by running `fit` for longer, but we do not do so here.

```{r}
#Load back the mapping contrast matrix, which may be used during simulation
load("img/rate_contrasts.RData")
#Load in the simulated data
load("data/simmed_group.RData")
# Load in the parameter estimates obtained from simulated data for both models
fitPM <- get(load("samples/samplerPM.RData"))
fitPMnc <- get(load("samples/samplerPM_noproactive.RData"))
#Get a summary of the pop mean parameters for the more complex model
mus <- summary(fitPM, selection="mu")
```

Ensure that parameter estimation converged successfully:
```{r eval=FALSE}

#The check function prints information about the samples and
# shows trace plots for the worst mu, sigma, and alpha parameter. Here we see that the A
# paramter is still changing slightly during the first half of the samples.
check(fitPM)
check(fitPMnc)

```


#### **Parameter Recovery** 
Here we check the recovery of population-level parameters by comparing estimates to the original simulated values.Both the population mean (mu) and individual level estimates (alpha) look reasonable, as indicated by a good fit
to the y=x line.

```{r}

#The recovery function returns a plot. For selection="mu", y is pop mean 
# parameter estimates and x is generating values. 
# Running print on recovery prints out a table containing the information
print(round(recovery(fitPM,layout=NULL,
               p_vector_sim,
               selection="mu"
)$mu$quantiles,2))


```

```{r}
#When given selection=alpha, plots examine estimates for each individual
# on Y axis, true generating values on x. Close to line y=x means a good fit
recovery(fitPM,
               res_recovery,
               selection="alpha"
)
#A similar printout is available to above, but not used here as output is large
```


#### **Model Recovery**
Another test is whether we can verify that the 'true' generating model structure is supported over alternative explanations, using model comparison tools. For an example, we compare the full model (which includes proactive-control effects) against an alternative model that excludes a proactive-control threshold effect used in the simulation.

We compare models using the Deviance Information Criterion (DIC) and Bayesian Predictive Information Criterion (BPIC). Lower values and higher weights indicate more relative support for a model.

```{r}
#The compare function looks at DIC/BPIC of a provided list of models.
# BayesFactor=F is set to avoid costly calculations of marginal likelihoods,
# but could be set to TRUE if Bayes Factor comparisons are desired (although this often requires getting a large number of samples).
compare(list(
  top = fitPM,
  noproactive = fitPMnc
  ), BayesFactor = F
)

```
 
 In this case, the full model outperforms the alternative, suggesting that proactive-control effects are important. Further tests could examine other the effect of excluding other mechanisms such as reactive control, as demonstrated in Lesson2.Rmd. 

