---
title: 'Lesson 2: Model generalization, unification and assessment'
author: "Dr. Luke Strickland"
date: "`r Sys.Date()`"
output:
  html_document: default
  pdf_document: default
  word_document: default
editor_options:
  markdown:
    wrap: 72
---

## Setup

We recommend downloading RStudio and loading this project by opening
ExpressingTheories_Tutorial.Rproj.

##### Install packages globally
The primary requirement for these scripts is the EMC2 package, 
which can be installed from CRAN with install.packages("EMC2").Tidyverse packages 
are also used for utility functions - specifically dplyr, tidyr, ggplot2, and 
stringr. These packages can all be installed from CRAN with install.packages().

- On Windows, some packages may require that Rtools is first installed.
- On macOS, some packages may require that command line developer tools are installed.

##### renv
Alternatively, we provide an renv environment, which you can use to automatically 
install all packages for this R project with the specific versions we used. These 
packages are kept isolated to this project and do not affect your global R library.
To use this rather than globally install packages, see the script setup_renv.R
If you have issues with renv, we recommend trying installing packages globally
instead.

### Pre-computed objects

MCMC samples require compute time. Posterior predictives also require some time, although far less. 
All pre-computed sample and posterior predict objects are available in samples.zip

Samples should be moved inside of the “samples” directory in order for the scripts to run.

## Overview

This script followed on from Lesson1.Rmd, which illustrated how to specify
and fit the Prospective Memory (PM) decision control (PMDC) model. Here,
we generalize these techniques to another paradigm. Specifically, we
illustrate an evidence accumulation model (EAM) quantifying
the mechanisms underlying automation use (Strickland et al., 2018). 
We demonstrate how to unify this model with the PMDC model reviewed in Lesson One.
We then walk through model assessment techniques applying the unified model
to an example data set. The unified model and associated data are from a recent
empirical study (Boag et al., 2025).

-   Boag, Strickland, Heathcote & Loft (2025) The Use of Automated Decision Aids in Multi-Tasking Context.*Under Review*.

-   Strickland, L., Heathcote, A., Bowden, V. K., Boag, R. J., Wilson, M. K., Khan, S., & Loft, S. (2021). Inhibitory cognitive control allows automated advice to improve accuracy while minimizing misuse. Psychological Science, 32(11), 1768-1781. https://doi.org/10.1177/09567976211012676

## Load

Load packages, bespoke utility functions and relevant data.

```{r results='hide', message=FALSE}
#Primary package for model fitting
library(EMC2)
#Data manipulation
library(dplyr)
#For plots
library(ggplot2)
#Some bespoke utility functions
source("utility_functions.R")

#Data and contrast matrix for automation example
load("data/AutoData.RData")
load("img/exinhcm_auto.RData")


#Data and contrast matrices for unified model
load("data/data_advanced.RData")
load("img/exinhcm_PM.RData")
load("img/Proactivecm_auto.RData")
```
    

# Automation Use Model
    
Strickland et al. (2021) investigated how people integrate automated advice into 
their decision-making.  
In their study, participants performed a conflict detection task,
where they determined whether 
aircraft pairs would violate safe separation standards during their flight. If so,
they were asked to classify aircraft as "conflict". If not, then "non-conflict".

Participants completed the task under two blocked conditions:

- **Manual condition:** participants made decisions unaided.  
- **Automation condition:** a decision aid (automation) provided a recommendation, 
either conflict (CONFLICT) or non-conflict (NON-CONF). The advice was correct
on 90% of automation-block trials. 

The data set is shown below.
It contains standard EMC2 columns, such as subjects,
R for response (C= conflict, N= non-conflict), and rt.

Also, design-specific columns:

-   `S`: stimulus type (c=conflict, n=nonconflict),
-   `auto`: Did participants have an automated decision aid in that
    block(auto) or not (manual)?
-   `failtrial`: Did the automated advice recommend a correct decision
    (AC) or incorrect (AI) on the particular trial?

Note that AC and AI are also specified for manual, but are just matched
trials with the factor having little meaning, as no automated advice was provided. 

```{r}
head(AutoData)
```    
    
In Strickland et al. (2018)'s model, automation-use affects accumulation rates in two  ways:

1) Excitation, where automation inputs increase accumulation towards
recommended responses. 
- *Excitation implies that accumulation rates
should be faster than manual rates for the choice automation agrees
with.*
2) Inhibition, where automation inputs decrease accumulation
towards opposing responses. 
- *Inhibition implies accumulation rates should be slower than manual for the choice automation disagrees with.*

Unlike the last example,
base ongoing-task accumulation rates are not specified in terms of "quality"
and "urgency" here, for simplicity. However, they could have been specified that way. 


# Mapping Contrasts

To express Strickland et al.’s mechanisms within EMC, we need 
contrasts that map estimated parameters (excitation/inhibition) onto 
design cell parameters.  
These contrasts specify how each accumulator’s evidence-accumulation rate depends on combinations of factors such as automation condition and advice congruence.

Automation-informed contrasts are added in automation blocks:
-  automation excitation is added to rates agreeing with automated advice (e.g., the matching accumulation rate when
automation is correct)
-  automation inhibition is subtracted from rates disagreeing with automated advice (e.g., the matching 
accumulation rate when automation is incorrect)


```{r}

#Print out the equations implied by the excitation-inhibition contrast matrix
for(i in rownames(exinhcm_auto)){
  print(read_rows(exinhcm_auto, i))
}
```

## Model Specification

We now demonstrate how to specify this model using design.

```{r results='hide'}

#Specifying accumulator/design cells where the accumulator is correct
match_fun <- function(d)
  as.character(substr(d$S,1,1)) == tolower(
    as.character(d$lR)
  )

design_auto <- design(
  #Same model type as Lesson One: LBA
  model=LBA,
  #Data which will be used to determine experiment factors and latent responses
  data=AutoData,
  # Defining a custom factor which will be used for accumulation-rate contrasts
  # Pastes together accumulator, stimulus type, auto vs manual condition,
  # whether automation is correct or incorrect
  functions=list(SlR=function(d) factor(paste0(d$lR, d$S, d$AM, d$fail),
                                   levels=rownames(exinhcm_auto))
                          ),
  #The contrast matrix determining the action of inhibition/excitation
  #(created in Lesson2_contrasts.R)
  contrasts = list(SlR=exinhcm_auto),
  matchfun=match_fun,
  #Key linear model formulae
  #v specified according to SlR and contrast matrix, thresholds can vary based on
  # accumulator and auto vs manual conditions, 1 A, t0. sv can vary by latent match
  formula=list(v~SlR,B ~ lR*AM,A ~ 1,t0 ~ 1,sv ~ lM),
    constants=c(
    # sv set to 1 here as a scaling parameter
    # To force it to be positive, EMC estimates log(sv) then exponentiates 
    # afterwards. Hence set to log(1)
    sv=log(1),
    # By default, an intercept parameter is also included, which we don't want 
    # as the contrasts already fully define the v parameters for this example. 
    #So set the intercept to constant 0
    v=0
  )
)
```

```{r}
#Giving mapped_pars just the design prints out the relevant linear models that are
# created. These models are used to determine design-cell parameters.
mapped_pars(design_auto)
```

# Unifying automation-use and PM models: Boag et al. (2025)

A major advantage of computational modelling is the ability to unify theories
across disparate domains. For example, both the PM and automation examples use 
feedforward control architectures to quantify how decisions are affected by inputs,
either PM stimuli or automated advice recommendations. With an appropriate design, 
these mechanisms can operate together within a single EAM. Boag et al. (2025) 
proposed a unified model including both. Here, we walk through their example
to illustrate model unification.

Boag et al. (2025) tested their model with an experiment combining a PM task
and automated advice. Participants performed
a conflict detection task. In some
conditions, participants had a PM task. This required pressing an alternative key
for targets with particular aircraft callsigns. In some conditions, participants were
provided with an automated decision aid that recommends whether
aircraft are in conflict or not. This decision aid was sometimes wrong. PM and 
automation conditions were crossed within-subjects for each participant.

**Boag et al. (2025) Design Factors**

-   *Stimulus Type*: Conflict, Nonconflict, or PM
-   *PM Block*: Control or PM
-   *Automation Block*: Manual (Unaided) or Automation (Aided)
-   *Automation Accuracy*: Correct or Incorrect

Below is the data set that will be analysed. It contains standard EMC2 columns, such as subjects,
R for response (C= conflict, N= non-conflict, P=PM), and rt.

Also, design-specific columns:

-   `S` for stimulus type (cc=conflict, nn=nonconflict, pc= PM target, conflict, pn= PM target, nonconflict)
-   `auto`: Did participants have an automated decision aid in that
    block(auto) or not (manual)?
-   `failtrial`: Did the automated advice recommend a correct decision
    (AC) or incorrect (AI) on the particular trial?
-   `PM`: PM or control conditions?

```{r}
head(dat)
```


## Contrasts

Having outlined the design, the next step is to formalize how the theoretical mechanisms
are represented.  
To instantiate PM and automation-use mechanisms, we specify a set of mapping contrasts 
that translate each experimental combination of PM and automation conditions into corresponding evidence-accumulation rates.  

As discussed in the manuscript, the idea is that the effects of automation and PM inputs sum. 
For example, an accumulation
rate could be increased by automation-excitation when it is congruent with automated advice,
but simultaneously reduced if inhibited due to the presence of PM target inputs: 

 - OTmatch, Automation Congruent, PM trial = v[OTmatch] + E_Auto - I_PM

The full set of mapping contrasts are demonstrated below.


```{r}
#The combined automation-PM contrast matrix, mapped to Boag et al.'s design
#Removing rows that doesn't practically arise in design for better print
print_full_v_cm <- full_v_cm[apply(full_v_cm, 1, function(row) any(row != 0)), ]
#Print each mapping contrast
for(i in rownames(print_full_v_cm )){
  print(read_rows(print_full_v_cm , i))
}
```

As with previous PMDC example, the possibility of proactive control over thresholds is included
(larger thresholds in PM than control blocks).
Furthermore, this proactive control is allowed to differ across conditions where participants
perform the task manually versus when they work with automated advice.

In this example, proactive control over thresholds will also be coded
using contrasts, rather than included as linear-model effects. It would also 
have been possible to use linear-model syntax to interact automation conditions, PM conditions,
and latent accumulator (lR).

```{r}

print_Proactivecm_auto <- Proactivecm_auto[
  apply(Proactivecm_auto, 1, function(row) any(row != 0)), 
  ]

for(i in rownames(print_Proactivecm_auto)){
  print(read_rows(print_Proactivecm_auto, i))
}
```

## Model Specification


We now demonstrate how to specify Boag et al. (2025)'s model and fit it to real data. 
We start by looking at the input to the design function. This is similar to the
first automation example, but requires the mapping contrasts specified above and different
custom factors. 

```{r results='hide'}

#Accuracy function: Checking whether accumulator matches stimulus type
match_fun <- function(d)
  as.character(substr(d$S,1,1)) == tolower(
    as.character(d$lR)
  )

design_advanced <- design(model=LBA, 
                          # The applied data set (Boag et al., 2025)
                          data=dat,
                          # The relevant factors for contrasts will match the 
                          # row names in the contrast matrices.
                          functions=list(
                          # An SlR factor here similar to previous examples,
                          # but including both automation and PM factors
                            SlR=function(d) factor(paste0(d$lR, d$S, d$AM, d$fail, d$PM),
                                                   levels=rownames(full_v_cm)),
                          # Also a PMlR factor added for the
                          # proactive control model
                            PMlR=function(d) factor(paste0(d$lR, d$PM, d$AM),
                                                    levels=rownames(Proactivecm_auto)),
                          # Special "RACE" factor which will be used to deal with
                          # the fact that PM has different number of accumulators to control
                            RACE=function(d)factor(ifelse(d$PM=="control",2,3))
                          ),
                          # Supplying the contrast matrices
                          contrasts = list(SlR=full_v_cm, PMlR =Proactivecm_auto),
                          matchfun=match_fun,
                          # Specifying how parameters vary by linear model formulae,
                          # similar to other examples
                          formula=list(v~SlR,B ~ PMlR,A ~ 1,t0 ~ 1,sv ~ lM),
                          # Appropriate constants
                          constants=c(sv=log(1),v=0, B=log(1))
)
```

Next, we set up a prior which will be used for parameter estimation. 
```{r}

#Setup a p_vector containing naive prior means
p_vector <- sampled_pars(design_advanced)
p_vector[grepl("B", names(p_vector))] <- log(1)
p_vector[grepl("inh", names(p_vector))] <- 0
p_vector[grepl("PMV", names(p_vector))] <- 1
p_vector[grepl("McC", names(p_vector))] <- 1
p_vector[grepl("MnN", names(p_vector))] <- 1
p_vector[grepl("t0", names(p_vector))] <- log(0.4)

n_vs <- length(p_vector[grepl("^v", names(p_vector))])
n_Bs <- length(p_vector[grepl("B", names(p_vector))])

#A vector of standard deviations to record uncertainty (larger sd means more uncertain)
psd=c(
  #drifts
  rep(2,n_vs),
  #Bs,
  rep(1,(n_Bs)),
  #A, t0, sv true
  1, 1, 1
)


```

Create EMC2's prior object. Then merge with data to make an emc sampler object,
and save off. Sampling is dispatched at the command line in another script (*run_fit.R*).
 
```{r eval=FALSE}
#Create prior object
prior_advanced  <- prior(design_advanced,mu_mean=p_vector, mu_sd=psd)
#Create emc sampler
emc <- make_emc(dat, design_advanced, prior=prior_advanced)
#save(emc,file="samples/sampler_advanced.RData")
```


#### **Sub-models**

We fit a range of candidate models that might
account for the advanced data set. 

-   "Top" flexible model with full flexibility in PM and automation
    mechanisms specified in the last contrasts scripts. Specified above

-   A model that excludes the PM-inhibition mechanism

-   A model that excludes the PM-induced proactive control mechanisms

-   A model that excludes the automation use mechanisms

These sub-model constraints are implemented by setting parameters to
a constant, but could equally be created by re-specifying contrasts

For brevity, comments outlining functionality that has already been explained
in the full model code chunk above are not repeated for the nested models.

##### **No PM inhibition**

No PM-inhibition mechanism for either auto or manual conditions. Set the PM inhibition parameters to 0.

```{r results='hide'}

# Similar arguments as the full model, except for constants.
design_advanced_noPMinh <- design(model=LBA, 
                                  data=dat,
                                  functions=list(
                                    SlR=function(d) factor(paste0(d$lR, d$S, d$AM, d$fail, d$PM),
                                                           levels=rownames(full_v_cm)),
                                    PMlR=function(d) factor(paste0(d$lR, d$PM, d$AM),
                                                            levels=rownames(Proactivecm_auto)),
                                    RACE=function(d)factor(ifelse(d$PM=="control",2,3))
                                  ),
                                  contrasts = list(SlR=full_v_cm, PMlR =Proactivecm_auto),
                                  matchfun=match_fun,
                                  formula=list(v~SlR,B ~ PMlR,A ~ 1,t0 ~ 1,sv ~ lM),
                                  constants=c(sv=log(1),v=0, B=log(1),
                                              #Fixing PM inhibition to 0.
                                              v_SlRMANPMinh=0,
                                              v_SlRAUTOPMinh=0
                                  )
)

#Creating a p vector with appropriate parameters for pop mean priors, based
# on previous p vector
p_vector_noPMinh <-
  p_vector[!names(p_vector) %in% c("v_SlRMANPMinh", "v_SlRAUTOPMinh")]

#Creating a psd with appropriate parameters, indexing uncertainty in pop means
n_vs_noPMinh <- length(p_vector_noPMinh[grepl("^v", names(p_vector_noPMinh))])
n_Bs_noPMinh <- length(p_vector_noPMinh[grepl("B", names(p_vector_noPMinh))])

psd_noPMinh =c(
  #drifts
  rep(2,n_vs_noPMinh),
  #Bs,
  rep(1,(n_Bs_noPMinh)),
  #A, t0, sv true
  1, 1, 1
)
```

```{r eval=FALSE}
prior_advanced_noPMinh  <- prior(design_advanced_noPMinh,
                                 mu_mean=p_vector_noPMinh, 
                                 mu_sd=psd_noPMinh
)
emc <- make_emc(dat, design_advanced_noPMinh, prior=prior_advanced_noPMinh)
#save(emc,file="samples/sampler_advanced_noPMinh.RData")
```


##### **No Proactive Control**

No proactive control. Set the proactive control parameters to log(1) (i.e., 0).

```{r results='hide'}
design_advanced_noproactive <- design(model=LBA, 
                                      data=dat,
                                      functions=list(
                                        SlR=function(d) factor(paste0(d$lR, d$S, d$AM, d$fail, d$PM),
                                                               levels=rownames(full_v_cm)),
                                        PMlR=function(d) factor(paste0(d$lR, d$PM, d$AM),
                                                                levels=rownames(Proactivecm_auto)),
                                        RACE=function(d)factor(ifelse(d$PM=="control",2,3))
                                      ),
                                      contrasts = list(SlR=full_v_cm, PMlR =Proactivecm_auto),
                                      matchfun=match_fun,
                                      formula=list(v~SlR,B ~ PMlR,A ~ 1,t0 ~ 1,sv ~ lM),
                                      constants=c(sv=log(1),v=0, B=log(1), 
                                                  #Fixing proactive control effects
                                                  # to 0
                                                  B_PMlRmanualpaC=log(1),
                                                  B_PMlRmanualpaN=log(1),
                                                  B_PMlRautopaC=log(1),
                                                  B_PMlRautopaN=log(1)
                                      )
)


p_vector_noproactive <-
  p_vector[!names(p_vector) %in% c("B_PMlRmanualpaC", "B_PMlRmanualpaN",
                                   "B_PMlRautopaC", "B_PMlRautopaN")]
n_vs_noproactive <- length(p_vector_noproactive[grepl("^v", names(p_vector_noproactive))])
n_Bs_noproactive <- length(p_vector_noproactive[grepl("B", names(p_vector_noproactive))])


psd_noproactive =c(
  #drifts
  rep(2,n_vs_noproactive),
  #Bs,
  rep(1,(n_Bs_noproactive)),
  #A, t0, sv true
  1, 1, 1
)
```

```{r eval=FALSE}
prior_advanced_noproactive  <- prior(design_advanced_noproactive,
                                     mu_mean=p_vector_noproactive, 
                                     mu_sd=psd_noproactive
)
emc <- make_emc(dat, design_advanced_noproactive, prior=prior_advanced_noproactive)

#save(emc,file="samples/sampler_advanced_noproactive.RData")
```

##### **No Automation Use Mechanisms**

Exclude automation-use mechanisms by fixing its parameters to 0.

```{r results='hide'}
design_advanced_noauto <- design(model=LBA, 
                                 data=dat,
                                 functions=list(
                                   SlR=function(d) factor(paste0(d$lR, d$S, d$AM, d$fail, d$PM),
                                                          levels=rownames(full_v_cm)),
                                   PMlR=function(d) factor(paste0(d$lR, d$PM, d$AM),
                                                           levels=rownames(Proactivecm_auto)),
                                   RACE=function(d)factor(ifelse(d$PM=="control",2,3))
                                 ),
                                 contrasts = list(SlR=full_v_cm, PMlR =Proactivecm_auto),
                                 matchfun=match_fun,
                                 formula=list(v~SlR,B ~ PMlR,A ~ 1,t0 ~ 1,sv ~ lM),
                                 constants=c(sv=log(1),v=0, B=log(1),
                                  #Fixing automation-use mechanisms to 0.           
                                             v_SlRPMAUTOex = 0, v_SlRPMAUTOinh =0,
                                             v_SlRcontrolAUTOex = 0, v_SlRcontrolAUTOinh =0
                                 )
)

p_vector_noauto <-
  p_vector[!names(p_vector) %in% c("v_SlRPMAUTOex", "v_SlRPMAUTOinh",
                                    "v_SlRcontrolAUTOex", "v_SlRcontrolAUTOinh")]

n_vs_noauto <- length(p_vector_noauto[grepl("^v", names(p_vector_noauto))])
n_Bs_noauto <- length(p_vector_noauto[grepl("B", names(p_vector_noauto))])

psd_noauto =c(
  #drifts
  rep(2,n_vs_noauto),
  #Bs,
  rep(1,(n_Bs_noauto)),
  #A, t0, sv true
  1, 1, 1
)
```


```{r eval=FALSE}
prior_advanced_noauto  <- prior(design_advanced_noauto,
                                mu_mean=p_vector_noauto, 
                                mu_sd=psd_noauto
)
emc <- make_emc(dat, design_advanced_noauto, prior=prior_advanced_noauto)

#save(emc,file="samples/sampler_advanced_noauto.RData")
```

## Model Assessment

Having setup and sampled all the previous models, we now use this as an opportunity
to illustrate model assessment in a complex example.

Model workflow often involves a break for fitting times, and so it is likely
this code might be run in a separate session. Code below to reload necessary
functions. Also, some tidyverse functions for summarizing data, predictions and plotting.

This custom code takes data and posterior predictives and
create summaries for fit plots. Although EMC2 also provides such functions, 
here we use illustrate how to 
apply some simple bespoke code to EMC2 outputs, giving results in 
line with Boag et al. (2025).

### Load

```{r}
#Packages and utility functions
library(tidyr)
library(dplyr)
library(ggplot2)
library(EMC2)
source("utility_functions.R")
#ggplot 2: set the theme for plots
theme_set(theme_bw())
#Contrast matrices used in the Boag et al. (2025) models, re-called during
# simulate
load("img/exinhcm_PM.RData")
load("img/Proactivecm_auto.RData")
```

Below loads samples from the most complex "top" model.
```{r}
fit_advanced <- get(load("samples/sampler_advanced.RData"))
```


#### **Evaluate Posterior Predictive Fit**

Get posterior predictives using EMC2's predict method and save. Commented out because
predict can take some time and be better
to dispatch multi-core. Predictions were provided in the samples folder on OSF.
```{r eval=FALSE}
#Check the model converged ok (it did)
check(fit_advanced)
#Get posterior predictives (somewhat slow)
# pp <- predict(fit_advanced, n_cores=1)
dat <- get_data(fit_advanced)
#save(pp, dat, file= "samples/predicted_advanced.RData")
```

Load up the prepared posterior predicitves.
```{r}
load("samples/predicted_advanced.RData")
```

EMC2 provides summary and plotting functions. However, is it also straightforward
to extract data and posterior predictives (as data frames) as a basis for specifying custom plots. 
Here, we aim to 
produce plots consistent with Boag et al. (2025). To that end, a bespoke utility function,
`calc_stat_postfit` takes `dat` and `pp` extracted from the EMC2 fit and summarizes
based on a provided function.

The custom function requires a data summary specified as function (here, accuracy),
experimental factors to examine, and `stat_cols` that are required for the function (here S and R 
are required to calculate accuracy). It returns summaries of the data and posterior predictives.
By default, the grouped data frame is examined, providing information about average fit. The
"subjects" column can also be applied to look at individual participants.

```{r}

# The bespoke calc_stat_postfit (from utility_functions.R) can calculate 
# any summary function and apply to data
# as well as each posterior predictive. Below do so for an accuracy function
acc_fun <- function(S, R){
  mean(substr(S,1,1)==tolower(R))
}

#Provide observed data (dat), posterior predictives (pp), factors to aggregate
# over (facs), the function to apply (stat_fn), and columns necessary to calculate
# the function (stat_cols)
acc_df <- calc_stat_postfit(dat, pp, facs=c("S", "AM", "fail", "PM"),
                            stat_fn=acc_fun,
                            stat_cols=c("S","R")
)

#Relabel a factor for nicer look on plots
acc_df$AM_fail <- factor(
  paste0(acc_df$AM, acc_df$fail),
  levels=c( "manualAC", "autoAC","autoAI"),
  labels=c("manual", "AC", "AI"))

```



Plots below visualize accuracy across
automation and failure conditions (`AM_fail`), split by stimulus type
(`S`) and PM condition.

-   The first plot focuses on `"cc"` and `"nn"` trials, faceted by `S`
    and `PM`.
-   The second plot excludes those trials and shows remaining `S` types
    side-by-side
-   Dashed lines connect grouped conditions (`nam`)
-   The black point layer (`stat`) provides comparison values (e.g.,
    model predictions or alternative metrics)

As plotted below, fits to both PM and ongoing-task accuracy look quite good. Fits are generally improved in 
the Boag et al. (2025) paper, which specified a more complex model.

```{r}
#dplyr pipe to filter out PM stimuli, followed by ggplot syntax to plot
#AM_fail specifies x axis and postmn_stat y
acc_df %>% filter(S %in% c("cc","nn")) %>% ggplot(aes(AM_fail, postmn_stat)) +
  #Add data as oen circles
  geom_point(size=3)+ 
  #Add model prediction points as filled circles (posterior mean)
  geom_point(aes(y= stat), pch=21, size=4, colour="black")+
  #Lines connecting model predictions
  geom_line(aes(group=1), linetype=2)+
  #Error bars for credible intervals
  geom_errorbar(aes(ymax = upper, ymin = lower), width=0.5) +
  #A grid where stimulus type is faceted vertically, and PM condition horizontally
  # Plus labels for axes
  facet_grid(S~PM) +xlab("Automation") +ylab("Accuracy")+
  #Some adjustments to text size and y axis range
  theme(text = element_text(size = 22), axis.text=element_text(size=15))+ylim(0.5,1)
```

```{r}
#Similar syntax to above, except filtering FOR PM trials. No control
# conditions needed here so changing up to facet by PM trial type (conf/nonconf) 
# rather than
# by PM/control
acc_df %>% filter(!S %in% c("cc","nn")) %>% ggplot(aes(AM_fail, postmn_stat)) +
  geom_point(size=3)+ geom_errorbar(aes(ymax = upper, ymin = lower), width=0.5) +
  geom_line(aes(group=1), linetype=2)+
  geom_point(aes(y= stat), pch=21, size=4, colour="black")+
  facet_grid(.~S) +xlab("Automation") +ylab("Accuracy")+
  theme(text = element_text(size = 22), axis.text=element_text(size=15)) +ylim(0.5,1)
```

For RT, we wish to examine three aspects of RT distribution: 0.1 quantile,
median (0.5 quantile), and 0.9 quantile. This gives an indication of distributional 
fit. Another non-EMC2 convenience function `stack_stats_postfits` can apply 
to get multiple summaries of data simultaneously and bind to a data frame.

Requires:

-   A list of functions (`corRT_fn_list`, defined earlier) to compute
    the quantiles
    `corRT_fn_list` computes the 10th,
50th, and 90th percentiles of reaction times for error trials
-   The columns to pass each function (`cols`). Equivalent to stat_cols in the earlier function
-   Names of the resulting statistics for the outputted data frame

```{r}

#The list of functions which will summarize data
corRT_fn_list <- c(function(S, R, rt){quantile(rt[toupper(substr(S,1,1))==R], probs=0.1, na.rm=T)},
                   function(S, R, rt){quantile(rt[toupper(substr(S,1,1))==R], probs=0.5, na.rm=T)},
                   function(S, R, rt){quantile(rt[toupper(substr(S,1,1))==R], probs=0.9, na.rm=T)}
)

#The columns required by the functions
cols <- c("S", "R", "rt")
# Supplied as a list, one entry for each function
stat_col_list <- list(cols,cols,cols)
#Names for outputs
stat_fn_names <- c("0.1", "0.5", "0.9")

#Function calc using data, pp, factors, functions, columns, and output names
cRT_df <- stack_stat_postfits(
  dat, pp, facs=c("S", "AM", "fail", "PM"), corRT_fn_list, stat_col_list, stat_fn_names
) 

#Relabel AM_fail for nicer plot
cRT_df$AM_fail <- factor(
  paste0(cRT_df$AM, cRT_df$fail),
  levels=c( "manualAC", "autoAC","autoAI"),
  labels=c("manual", "AC", "AI"))
```

Plots below show good distributional fits to correct RT, other than a small miss
on the 0.9 quantile for the rare PM response times.

```{r}
#Similar dplyr/ggplot syntax as previous examples, except there are now
# lines connecting points for each quantile
#Starting with non-PM trials
cRT_df %>% filter(S %in% c("cc","nn")) %>%  ggplot(aes(AM_fail, postmn_stat)) +
  geom_point(size=3)+ geom_errorbar(aes(ymax = upper, ymin = lower), width=0.5) +
  #Lines below connect group by quantiles ("nam" generic output of the function above)
  geom_line(aes(group=nam), linetype=2)+
  geom_point(aes(y= stat), pch=21, size=4, colour="black")+
  facet_grid(S~PM)  +ylab("Correct RT")  +xlab("Automation")+
  theme(text = element_text(size = 22), axis.text=element_text(size=15))

#Followed by PM trials
cRT_df %>% filter(!S %in% c("cc","nn")) %>%  ggplot(aes(AM_fail, postmn_stat)) +
  geom_point(size=3)+ geom_errorbar(aes(ymax = upper, ymin = lower), width=0.5) +
  geom_line(aes(group=nam), linetype=2)+
  geom_point(aes(y= stat), pch=21, size=4, colour="black")+
  facet_grid(.~S)  +ylab("Correct RT")  +xlab("Automation")+
  theme(text = element_text(size = 22), axis.text=element_text(size=15))
```


Error RTs visualized similarly below, except for PM trials due to this being a 
rare type of response. 

```{r}
errRT_fn_list <- c(function(S, R, rt){quantile(rt[toupper(substr(S,1,1))!=R],
                                               probs=0.1)},
                   function(S, R, rt){quantile(rt[toupper(substr(S,1,1))!=R], 
                                               probs=0.5)},
                   function(S, R, rt){quantile(rt[toupper(substr(S,1,1))!=R], 
                                               probs=0.9)}
)

eRT_df <- stack_stat_postfits(
  dat, pp, facs=c("S", "AM", "fail", "PM"), errRT_fn_list, stat_col_list, stat_fn_names
)

eRT_df$AM_fail <- factor(
  paste0(eRT_df$AM, eRT_df$fail),
  levels=c( "manualAC", "autoAC","autoAI"),
  labels=c("manual", "AC", "AI"))


eRT_df %>% filter(S %in% c("cc","nn")) %>% ggplot(aes(AM_fail, postmn_stat)) + 
  geom_point(size=3)+ geom_errorbar(aes(ymax = upper, ymin = lower), width=0.5) +
  geom_line(aes(group=nam), linetype=2)+
  geom_point(aes(y= stat), pch=21, size=4, colour="black")+
  facet_grid(S~PM)  +ylab("Error RT") +
  theme(text = element_text(size = 22), axis.text=element_text(size=15))
```

#### **Model Comparison**

After determining that the top model can fit the data, here we test whether 
the model structure can be simplified. Smaller DIC/BPIC supports selection a model. 
DIC/BPIC also help to determine the relative contributions of different
mechanisms to overall fit - larger differences in favour of a model suggest a stronger preference.


```{r}
#Load alternative models with mechanisms excluded
fit_advanced_noauto <- 
  get(load("samples/sampler_advanced_noauto.RData"))
fit_advanced_noPMinh <- 
  get(load("samples/sampler_advanced_noPMinh.RData"))
fit_advanced_noproactive <- 
  get(load("samples/sampler_advanced_noproactive.RData"))
```

All submodels converged in terms of Rhat <1.1 all parameters.
```{r eval=FALSE}
check(fit_advanced_noauto)
check(fit_advanced_noPMinh)
check(fit_advanced_noproactive)
```


Below we use EMC2's compare function to examine whether dropping mechanisms is supported.
Both DIC and BPIC suggest that the top model, with full PM and
automation-use mechanisms, is most supported. Thus, all the
included mechanisms are useful for predicting the data.

Bayes Factors could also have been calculated *by setting BayesFactor=TRUE), but they would require 
more posterior samples and were not used by Boag et al. (2025). 

```{r}
#Usage of EMC2's compare function:
# A list of samples to be compared, and specifying that no Bayes Factors
# should be calculated
out <- compare(list(
  top = fit_advanced,
  noPMinh = fit_advanced_noPMinh,
  noproactive = fit_advanced_noproactive,
  noauto = fit_advanced_noauto
), BayesFactor = FALSE)

#Rebase the DIC and BPIC values so that the minimum is 0
out$DIC <- out$DIC - out$DIC[1]
out$BPIC <- out$BPIC - out$BPIC[1]
out
```

#### **Model Parameter Summaries**

Having selected a most-preferred model, summarise the parameters to understand the underlying mechanisms implied.

1.  Extract population mean parameters from the full model
2.  Plot the posterior means/ 95% credible intervals
3.  Examine Bayesian p-values and Savage-Dickey Bayes Factors to evaluate effects

EMC2's get_pars function extracts parameters, here we focus on 'mu' (population means).
```{r}
#Give get_pars the required samples, ask for mu, merge_chains puts everything
# together rather than separating the samples for the different chains run
# during sampling
params <- data.frame(
  get_pars(fit_advanced, selection="mu", merge_chains = T)$mu[[1]]
)
```

##### Excitation and Inhibition

First, examine the excitation and inhibition mechanisms, both involving
automation and the PM task. 

This reveals:
-  strong PM inhibition in both auto and manual conditions
-  strong automation inhibition in the PM condition, but less in control
-  excitation strategy strong in control conditions, but not PM condition

```{r}
# Use some dplyr and tidyr here to summarize samples. First take the samples
# to a long-form data frame with pivot longer, each parameter name
# being defined in a column. Then summarize for each parameter posterior mean
# and 95% credible intervals.
param_summaries <- params %>% pivot_longer(cols=colnames(params)) %>% 
  group_by(name) %>% 
  summarize(mn=mean(value), 
            p025= quantile(value, probs=0.025),
            p975=quantile(value, probs=0.975)
  )

#Select the subset of excitation and inhibition parameters
exinh <- param_summaries  %>% filter(grepl(".*[ex|inh]$", name))

#Give them nice names
exinh$name <- c("PM inhibition, auto condition", 
                "PM inhibition, manual condition",
                "Automation excitation, PM condition",
                "Automation inhibition, PM condition",
                "Automation excitation, control condition",
                "Automation inhibition, control condition"
)

#Plot them
exinh  %>% 
  #Here putting the parameter names on x axis and values on y, but will flip at end
  ggplot(aes(x=name, y=mn, ymin=p025, ymax=p975))+
  ylab("Effect")+ xlab("") +
  #Geom_pointrange for the type of plot we are after (except we will flip it)
  geom_pointrange()+ 
  geom_hline(yintercept=0, lty=2) +  # add a dotted line at x=1 after flip
  #Do the flip and adjust text
  coord_flip() +
  theme(text = element_text(size = 18))


```

Having visualized the excitation and inhibition parameters, now evaluate the
strength and support for the effects and compare them. We do so by calculating 
Bayesian p-value (posterior sample density <0) and 
Savage-Dickey test (Bayes Factor of alternative vs null of 0) using EMC2's
`credible` and `hypothesis` functions. 

```{r results='hide', fig.show = "hide"}

#set.seed so that the Bayes Factor approximations are always the same,
# as they involve sampling from the prior
set.seed(1234)

#Use EMC2's credible function to get Bayesian p-value. Supply samples,
# the name of the parameter.
PMInh_auto_p <- attr(
  credible(x = fit_advanced, x_name = "v_SlRAUTOPMinh"), "less"
)

#Use EMC2's hypothesis function to get Savage-Dickey Bayes Factor. Supply samples,
# the name of the parameter.
PMInh_auto_SD <- hypothesis(fit_advanced, parameter="v_SlRAUTOPMinh",  do_plot = F)

#Same for other parameters below.
PMInh_man_p <- attr(
  credible(x = fit_advanced, x_name = "v_SlRMANPMinh"), "less"
)
PMInh_man_SD <- hypothesis(fit_advanced, parameter="v_SlRMANPMinh",  do_plot = F)


AutoInh_PM_p <- attr(
  credible(x = fit_advanced, x_name = "v_SlRPMAUTOinh"), "less"
)
AutoInh_PM_SD <- hypothesis(fit_advanced, parameter="v_SlRPMAUTOinh",  do_plot = F)

AutoInh_control_p <- attr(
  credible(x = fit_advanced, x_name = "v_SlRcontrolAUTOinh"), "less"
)
AutoInh_control_SD <- hypothesis(fit_advanced, parameter="v_SlRcontrolAUTOinh",  do_plot = F)


AutoEx_PM_p <- attr(
  credible(x = fit_advanced, x_name = "v_SlRPMAUTOex"), "less"
)
AutoEx_PM_SD <- hypothesis(fit_advanced, parameter="v_SlRPMAUTOex",  do_plot = F)


AutoEx_control_p <- attr(
  credible(x = fit_advanced, x_name = "v_SlRcontrolAUTOex"), "less"
)
AutoEx_control_SD <- hypothesis(fit_advanced, parameter="v_SlRcontrolAUTOex",  do_plot = F)

#Now look at differences in parameters.
#For credible, supply a function (x_fun) rather than an x name to look at differences
PM_control_AutoInh_p <- attr(
  credible(x = fit_advanced, x_fun = function(d) d["v_SlRPMAUTOinh"]
         - d["v_SlRcontrolAUTOinh"]),
  "less"
)

#Similarly, for hypothesis, provide a function (fun) rather than a parameter
PM_control_AutoInh_SD <- hypothesis(
  fit_advanced, fun=function(d) d["v_SlRPMAUTOinh"]
           - d["v_SlRcontrolAUTOinh"], do_plot = F)

PM_control_AutoEx_p <- attr(
  credible(x = fit_advanced, x_fun = function(d) d["v_SlRcontrolAUTOex"] 
         - d["v_SlRPMAUTOex"]), "less"
)

PM_control_AutoEx_SD <- hypothesis(fit_advanced, fun= function(d) 
  exp(d["v_SlRcontrolAUTOex"]) - exp(d["v_SlRPMAUTOex"]), do_plot = F)
```


```{r}
#Put together some nice names for the things we just calculated

effnames <- c("PM inhibition, manual", "PM inhibition, auto",
              "Auto inhibition, PM", "Auto inhibition, control",
              "Auto excitation, PM", "Auto excitation, control",
              "Auto inhibition, PM - Auto inhibition, control",
              "Auto excitation, PM - Auto excitation, control")

#Bind together the Bayesian p-values and Savage-Dickey Bayes Factors
ps <- c(PMInh_auto_p, PMInh_man_p,AutoInh_PM_p, AutoInh_control_p,
        AutoEx_PM_p, AutoEx_control_p, PM_control_AutoInh_p, 
        PM_control_AutoEx_p)

SDs <- c(PMInh_auto_SD, PMInh_man_SD,AutoInh_PM_SD, AutoInh_control_SD,
        AutoEx_PM_SD, AutoEx_control_SD, PM_control_AutoInh_SD, 
        PM_control_AutoEx_SD)

#Place everything in a data frame for a quick look and print it
df <- data.frame(Effect=effnames, p=round(ps, 3), SD= round(SDs, 3))

df

```


##### Proactive Control

Now we examine proactive control using similar techniques to above. 
This reveals:
- strong proactive control of both accumulators in both conditions
- weak tendency to less control in automation conditions

```{r}
proactive <- param_summaries %>% filter(grepl(".*pa.$", name))

proactive$name <- c("Conflict Accumulator, Automation", 
                    "Non-conflict Accumulator, Automation", 
                    "Conflict Accumulator, Manual", 
                    "Non-conflict Accumulator, Manual" 
)

proactive  %>% 
  ggplot(aes(x=name, y=mn, ymin=p025, ymax=p975))+
  ylab("Proactive Control Effect")+ xlab("") +
  geom_pointrange()+ 
  geom_hline(yintercept=0, lty=2) +  # add a dotted line at x=1 after flip
  coord_flip() +theme(text=element_text(size=14)) 

```

Bayesian p values and SD tests.
```{r results="hide", fig.show = "hide"}

set.seed(1234)
ProAutoC_p <- attr(credible(x = fit_advanced, x_name= "B_PMlRautopaC"),
                 "less")
ProAutoC_SD <- hypothesis(fit_advanced, parameter="B_PMlRautopaC",  do_plot = F)


ProAutoN_p <- attr(credible(x = fit_advanced, x_name= "B_PMlRautopaN"),
                 "less")
ProAutoN_SD <- hypothesis(fit_advanced, parameter="B_PMlRautopaN",  do_plot = F)



ProManC_p <- attr(credible(x = fit_advanced, x_name= "B_PMlRmanualpaC"),
                 "less")
ProManC_SD <- hypothesis(fit_advanced, parameter="B_PMlRmanualpaC",  do_plot = F)


ProManN_p <- attr(credible(x = fit_advanced, x_name= "B_PMlRmanualpaN"),
                 "less")
ProManN_SD <- hypothesis(fit_advanced, parameter="B_PMlRmanualpaN",  do_plot = F)



ProManAutoC_p <-  attr(credible(x = fit_advanced, x_fun = function(d) d["B_PMlRmanualpaC"] - d["B_PMlRautopaC"]
         ), "less"
)

ProManAutoC_SD <-  hypothesis(fit_advanced, fun = function(d) d["B_PMlRmanualpaC"] - 
             d["B_PMlRautopaC"],  do_plot = F
)



ProManAutoN_p <-  attr(credible(x = fit_advanced, x_fun = function(d) d["B_PMlRmanualpaN"] - d["B_PMlRautopaN"]
         ), "less"
)

ProManAutoN_SD <-  hypothesis(fit_advanced, fun = function(d) d["B_PMlRmanualpaN"] - 
             d["B_PMlRautopaN"],  do_plot = F
)


```

Simple proactive-control effects are large, but note the interaction (manual - auto)
is quite weak, with only a tendency for increase in manual not supported by p value
or SD test.
```{r}

effnames <- c("Conflict, Automation", 
              "Non-conflict, Automation", 
               "Conflict, Manual", 
              "Non-conflict, Manual",
              "Conflict, Manual - Conflict, Automation",
              "Non-conflict, Manual - Non-conflict, Automation")

ps <- 
  c(ProAutoC_p, ProAutoN_p, ProManC_p, ProManN_p, ProManAutoC_p,ProManAutoN_p)

SDs <- 
  c(ProAutoC_SD, ProAutoN_SD, ProManC_SD, ProManN_SD, ProManAutoC_SD,ProManAutoN_SD)

df <- data.frame(Effect=effnames, p=round(ps, 3), SD= round(SDs, 3))

df

```

As discussed in text, the proactive control calculated above was on the 
estimated scale, rather than on mapped parameters. As thresholds are mapped from the
log scale, this implies multiplicative rather than additive effects. To examine
proactive control on the natural scale, we can use "mapped" parameters and then
calculate proactive control effects post-hoc. Below, we show how to do so.

New proactive control plot and new p-values (~0.01, ~0.02 for conflict and non-conflict,
respectively).

```{r}
set.seed(1234)

#Extract mapped parameter values here, rather than estimated values. The key
# argument is map=T
params_mapped <- data.frame(
  get_pars(fit_advanced, selection="mu", merge_chains = T, map=T)$mu[[1]]
)

#Calculate proactive control effects from the mapped estimates for both
# auto and manual conditions
params_mapped$ProAC <- params_mapped$B_PMlRCPMauto - params_mapped$B_PMlRCcontrolauto 
params_mapped$ProAN <- params_mapped$B_PMlRNPMauto - params_mapped$B_PMlRNcontrolauto 

params_mapped$ProMC <- params_mapped$B_PMlRCPMmanual - params_mapped$B_PMlRCcontrolmanual 
params_mapped$ProMN <- params_mapped$B_PMlRNPMmanual - params_mapped$B_PMlRNcontrolmanual

#Calculate the differences in proactive control across manual and auto conditions
params_mapped$ProMAC <- params_mapped$ProMC - params_mapped$ProAC 
params_mapped$ProMAN <- params_mapped$ProMN - params_mapped$ProAN 


params_mapped_summaries <- params_mapped %>% pivot_longer(cols=colnames(params_mapped )) %>% 
  group_by(name) %>% 
  summarize(mn=mean(value), 
            p025= quantile(value, probs=0.025),
            p975=quantile(value, probs=0.975)
  )


proactive <- params_mapped_summaries %>% filter(grepl("Pro", name) & !(name %in% c("ProMAC", "ProMAN")))

proactive$name <- c("Conflict Accumulator, Automation", 
                    "Non-conflict Accumulator, Automation", 
                    "Conflict Accumulator, Manual", 
                    "Non-conflict Accumulator, Manual" 
)

# Strong proactive control in both auto and manual conditions. Slightly
# stronger for manual, perhaps due to strong automation-inhibition effects 
# providing support in PM conditions

proactive  %>% 
  ggplot(aes(x=name, y=mn, ymin=p025, ymax=p975))+
  ylab("Proactive Control Effect (Additive)")+ xlab("") +
  geom_pointrange()+ 
  geom_hline(yintercept=0, lty=2) +  # add a dotted line at x=1 after flip
  coord_flip() 


#Get p-values and Bayes Factors for the mapped, transformed values
# This involves similar calls to credible and hypothesis as earlier, except
# we are supplying mapping functions and subsequent post-map control calculations
# using the function arguments
additive_ProManAutoC_p <-  attr(credible(x = fit_advanced, x_fun = function(d) (
  (exp(d["B_PMlRmanualC"] +d["B_PMlRmanualpaC"])- exp(d["B_PMlRmanualC"]))  
  - (exp(d["B_PMlRautoC"] +d["B_PMlRautopaC"]) -exp(d["B_PMlRautoC"]))
  )
), "less"
)

additive_ProManAutoC_SD <-  hypothesis(fit_advanced, fun = function(d) (
 
  (exp(d["B_PMlRmanualC"] +d["B_PMlRmanualpaC"])- exp(d["B_PMlRmanualC"]))  
  - (exp(d["B_PMlRautoC"] +d["B_PMlRautopaC"]) -exp(d["B_PMlRautoC"]))
  ),  do_plot = F
)

additive_ProManAutoC_SD 
  
additive_ProManAutoN_p <-  attr(credible(x = fit_advanced, x_fun = function(d) (
 
  (exp(d["B_PMlRmanualN"] +d["B_PMlRmanualpaN"])- exp(d["B_PMlRmanualN"]))  
  - (exp(d["B_PMlRautoN"] +d["B_PMlRautopaN"]) -exp(d["B_PMlRautoN"]))
  )
), "less"
)

additive_ProManAutoN_SD <-  hypothesis(fit_advanced, fun = function(d) (
 
  (exp(d["B_PMlRmanualN"] +d["B_PMlRmanualpaN"])- exp(d["B_PMlRmanualN"]))  
  - (exp(d["B_PMlRautoN"] +d["B_PMlRautopaN"]) -exp(d["B_PMlRautoN"]))
  ),  do_plot = F
)

additive_ProManAutoN_SD


```
