---
title: 'Lesson 2: Supplementary, bounded parameters'
author: "Dr. Luke Strickland"
date: "`r Sys.Date()`"
output:
  pdf_document: default
  html_document: default
editor_options:
  markdown:
    wrap: 72
---

remotes::install_github("ampl-psych/EMC2", dependencies=TRUE, ref = "dev")

## Overview

Lesson 2 but with parameter bounds on excitation/inhibition introduced with a
pre-transform. See Lesson2.Rmd for more details.

## Load

Load packages, bespoke utility functions and relevant data.

```{r results='hide', message=FALSE}
#Primary package for model fitting
library(EMC2)
#Data manipulation
library(dplyr)
#For plots
library(ggplot2)
#Some bespoke utility functions
source("utility_functions.R")

load("data/data_advanced.RData")
load("img/exinhcm_PM.RData")
load("img/Proactivecm_auto.RData")
```


## Model Specification

Bounds implemented with pre-transforms.

Full model:

```{r results='hide'}

match_fun <- function(d)
  as.character(substr(d$S,1,1)) == tolower(
    as.character(d$lR)
  )

design_advanced <- design(model=LBA, 
                          data=dat,
                          functions=list(
                            SlR=function(d) factor(paste0(d$lR, d$S, d$AM, d$fail, d$PM),
                                                   levels=rownames(full_v_cm)),
                            PMlR=function(d) factor(paste0(d$lR, d$PM, d$AM),
                                                    levels=rownames(Proactivecm_auto)),
                            RACE=function(d)factor(ifelse(d$PM=="control",2,3))
                          ),
                          contrasts = list(SlR=full_v_cm, PMlR =Proactivecm_auto),
                          matchfun=match_fun,
                          formula=list(v~SlR,B ~ PMlR,A ~ 1,t0 ~ 1,sv ~ lM),
                          constants=c(sv=log(1),v=0, B=log(1)
                          ),
                          #Bounds implemented by putting "exp" pre-transform on
                          # excitation and inhibition parameters.
                          pre_transform = list(func = c(v_SlRcontrolAUTOex = "exp",
                            v_SlRcontrolAUTOinh = "exp",v_SlRPMAUTOex = "exp",v_SlRPMAUTOinh = "exp",
                            v_SlRAUTOPMinh = "exp", v_SlRMANPMinh = "exp"))
)
```

```{r}
p_vector <- sampled_pars(design_advanced)
p_vector[grepl("B", names(p_vector))] <- log(1)
#Prior inh 0 i.e., log(1)
p_vector[grepl("inh", names(p_vector))] <- 0
p_vector[grepl("PMV", names(p_vector))] <- 1
p_vector[grepl("McC", names(p_vector))] <- 1
p_vector[grepl("MnN", names(p_vector))] <- 1
p_vector[grepl("t0", names(p_vector))] <- log(0.4)

n_vs <- length(p_vector[grepl("^v", names(p_vector))])
n_Bs <- length(p_vector[grepl("B", names(p_vector))])

psd=c(
  #drifts
  rep(2,n_vs),
  #Bs,
  rep(1,(n_Bs)),
  #A, t0, sv true
  1, 1, 1
)


```

```{r eval=FALSE}
prior_advanced  <- prior(design_advanced,mu_mean=p_vector, mu_sd=psd)
emc <- make_emc(dat, design_advanced, prior=prior_advanced)
#save(emc,file="samples/sampler_advanced_bound.RData")
```


#### **Sub-models**

##### **No PM inhibition**

```{r results='hide'}
design_advanced_noPMinh <- design(model=LBA, 
                                  data=dat,
                                  functions=list(
                                    SlR=function(d) factor(paste0(d$lR, d$S, d$AM, d$fail, d$PM),
                                                           levels=rownames(full_v_cm)),
                                    PMlR=function(d) factor(paste0(d$lR, d$PM, d$AM),
                                                            levels=rownames(Proactivecm_auto)),
                                    RACE=function(d)factor(ifelse(d$PM=="control",2,3))
                                  ),
                                  contrasts = list(SlR=full_v_cm, PMlR =Proactivecm_auto),
                                  matchfun=match_fun,
                                  formula=list(v~SlR,B ~ PMlR,A ~ 1,t0 ~ 1,sv ~ lM),
                                  constants=c(sv=log(1),v=0, B=log(1),
                                              v_SlRMANPMinh=0,
                                              v_SlRAUTOPMinh=0
                                  ),
                          pre_transform = list(func = c(v_SlRcontrolAUTOex = "exp",
                            v_SlRcontrolAUTOinh = "exp",v_SlRPMAUTOex = "exp",v_SlRPMAUTOinh = "exp"))
)

p_vector_noPMinh <-
  p_vector[!names(p_vector) %in% c("v_SlRMANPMinh", "v_SlRAUTOPMinh")]
n_vs_noPMinh <- length(p_vector_noPMinh[grepl("^v", names(p_vector_noPMinh))])
n_Bs_noPMinh <- length(p_vector_noPMinh[grepl("B", names(p_vector_noPMinh))])

psd_noPMinh =c(
  #drifts
  rep(2,n_vs_noPMinh),
  #Bs,
  rep(1,(n_Bs_noPMinh)),
  #A, t0, sv true
  1, 1, 1
)
```

```{r eval=FALSE}
prior_advanced_noPMinh  <- prior(design_advanced_noPMinh,
                                 mu_mean=p_vector_noPMinh, 
                                 mu_sd=psd_noPMinh
)
emc <- make_emc(dat, design_advanced_noPMinh, prior=prior_advanced_noPMinh)
#save(emc,file="samples/sampler_advanced_bound_noPMinh.RData")
```

##### **No Proactive Control**

```{r results='hide'}
design_advanced_noproactive <- design(model=LBA, 
                                      data=dat,
                                      functions=list(
                                        SlR=function(d) factor(paste0(d$lR, d$S, d$AM, d$fail, d$PM),
                                                               levels=rownames(full_v_cm)),
                                        PMlR=function(d) factor(paste0(d$lR, d$PM, d$AM),
                                                                levels=rownames(Proactivecm_auto)),
                                        RACE=function(d)factor(ifelse(d$PM=="control",2,3))
                                      ),
                                      contrasts = list(SlR=full_v_cm, PMlR =Proactivecm_auto),
                                      matchfun=match_fun,
                                      formula=list(v~SlR,B ~ PMlR,A ~ 1,t0 ~ 1,sv ~ lM),
                                      constants=c(sv=log(1),v=0, B=log(1), 
                                                  B_PMlRmanualpaC=log(1),
                                                  B_PMlRmanualpaN=log(1),
                                                  B_PMlRautopaC=log(1),
                                                  B_PMlRautopaN=log(1)
                                      ),
                          pre_transform = list(func = c(v_SlRcontrolAUTOex = "exp",
                            v_SlRcontrolAUTOinh = "exp",v_SlRPMAUTOex = "exp",v_SlRPMAUTOinh = "exp",
                            v_SlRAUTOPMinh = "exp", v_SlRMANPMinh = "exp"))
)


p_vector_noproactive <-
  p_vector[!names(p_vector) %in% c("B_PMlRmanualpaC", "B_PMlRmanualpaN",
                                   "B_PMlRautopaC", "B_PMlRautopaN")]
n_vs_noproactive <- length(p_vector_noproactive[grepl("^v", names(p_vector_noproactive))])
n_Bs_noproactive <- length(p_vector_noproactive[grepl("B", names(p_vector_noproactive))])


psd_noproactive =c(
  #drifts
  rep(2,n_vs_noproactive),
  #Bs,
  rep(1,(n_Bs_noproactive)),
  #A, t0, sv true
  1, 1, 1
)
```

```{r eval=FALSE}
prior_advanced_noproactive  <- prior(design_advanced_noproactive,
                                     mu_mean=p_vector_noproactive, 
                                     mu_sd=psd_noproactive
)
emc <- make_emc(dat, design_advanced_noproactive, prior=prior_advanced_noproactive)

#save(emc,file="samples/sampler_advanced_bound_noproactive.RData")
```

##### **No Automation Use Mechanisms**

```{r results='hide'}
design_advanced_noauto <- design(model=LBA, 
                                 data=dat,
                                 functions=list(
                                   SlR=function(d) factor(paste0(d$lR, d$S, d$AM, d$fail, d$PM),
                                                          levels=rownames(full_v_cm)),
                                   PMlR=function(d) factor(paste0(d$lR, d$PM, d$AM),
                                                           levels=rownames(Proactivecm_auto)),
                                   RACE=function(d)factor(ifelse(d$PM=="control",2,3))
                                 ),
                                 contrasts = list(SlR=full_v_cm, PMlR =Proactivecm_auto),
                                 matchfun=match_fun,
                                 formula=list(v~SlR,B ~ PMlR,A ~ 1,t0 ~ 1,sv ~ lM),
                                 constants=c(sv=log(1),v=0, B=log(1),
                                             v_SlRPMAUTOex = 0, v_SlRPMAUTOinh =0,
                                             v_SlRcontrolAUTOex = 0, v_SlRcontrolAUTOinh =0
                                 ),
                          pre_transform = list(func = c(
                            v_SlRAUTOPMinh = "exp", v_SlRMANPMinh = "exp"))
)

p_vector_noauto <-
  p_vector[!names(p_vector) %in% c("v_SlRPMAUTOex", "v_SlRPMAUTOinh",
                                    "v_SlRcontrolAUTOex", "v_SlRcontrolAUTOinh")]

n_vs_noauto <- length(p_vector_noauto[grepl("^v", names(p_vector_noauto))])
n_Bs_noauto <- length(p_vector_noauto[grepl("B", names(p_vector_noauto))])

psd_noauto =c(
  #drifts
  rep(2,n_vs_noauto),
  #Bs,
  rep(1,(n_Bs_noauto)),
  #A, t0, sv true
  1, 1, 1
)
```

```{r eval=FALSE}
prior_advanced_noauto  <- prior(design_advanced_noauto,
                                mu_mean=p_vector_noauto, 
                                mu_sd=psd_noauto
)
emc <- make_emc(dat, design_advanced_noauto, prior=prior_advanced_noauto)

#save(emc,file="samples/sampler_advanced_bound_noauto.RData")
```

##### **No Automation-Excitation, PM condition**

No excitation from automation specifically in the PM condition. Parameter estimates 
suggested fitting this model - low population-mean parameter (~0).

```{r results='hide'}

design_noautoexPM <- design(model=LBA, 
                          data=dat,
                          functions=list(
                            SlR=function(d) factor(paste0(d$lR, d$S, d$AM, d$fail, d$PM),
                                                   levels=rownames(full_v_cm)),
                            PMlR=function(d) factor(paste0(d$lR, d$PM, d$AM),
                                                    levels=rownames(Proactivecm_auto)),
                            RACE=function(d)factor(ifelse(d$PM=="control",2,3))
                          ),
                          contrasts = list(SlR=full_v_cm, PMlR =Proactivecm_auto),
                          matchfun=match_fun,
                          formula=list(v~SlR,B ~ PMlR,A ~ 1,t0 ~ 1,sv ~ lM),
                          constants=c(sv=log(1),v=0, B=log(1),
                                             v_SlRPMAUTOex = 0
                          ),
                          pre_transform = list(func = c(v_SlRcontrolAUTOex = "exp",
                            v_SlRcontrolAUTOinh = "exp",v_SlRPMAUTOinh = "exp",
                            v_SlRAUTOPMinh = "exp", v_SlRMANPMinh = "exp"))
)
```

```{r}
p_vector_noautoexPM <-
  p_vector[!names(p_vector) %in% c("v_SlRPMAUTOex")]
n_vs_noautoexPM <- length(p_vector_noautoexPM[grepl("^v", names(p_vector_noautoexPM))])
n_Bs_noautoexPM <- length(p_vector_noautoexPM[grepl("B", names(p_vector_noautoexPM))])

psd_noautoexPM =c(
  #drifts
  rep(2,n_vs_noautoexPM ),
  #Bs,
  rep(1,(n_Bs_noautoexPM)),
  #A, t0, sv true
  1, 1, 1
)


```


```{r eval=FALSE}
prior_noautoexPM  <- prior(design_noautoexPM,mu_mean=p_vector_noautoexPM, mu_sd=psd_noautoexPM)
emc <- make_emc(dat, design_noautoexPM, prior=prior_noautoexPM)
# save(emc,file="samples/sampler_noautoexPMa.RData")
```


## Model Evaluation


```{r}
library(tidyr)
library(dplyr)
library(ggplot2)
library(EMC2)
theme_set(theme_bw())
source("utility_functions.R")
load("img/exinhcm_PM.RData")
load("img/Proactivecm_auto.RData")
```

```{r}
fit_advanced <- get(load("samples/sampler_advanced_bound.RData"))
```


#### **Evaluate Posterior Predictive Fit**

Predictives specifically for the bounded case.
```{r eval=FALSE}
check(fit_advanced)
pp <- predict(fit_advanced, n_cores=1)
dat <- get_data(fit_advanced)
#save(pp, dat, file= "samples/predicted_advanced_bound.RData")
```


```{r}
load("samples/predicted_advanced_bound.RData")
```

Fits with the theoretically plausible bounds look good.

```{r}
acc_fun <- function(S, R){
  mean(substr(S,1,1)==tolower(R))
}

acc_df <- calc_stat_postfit(dat, pp, facs=c("S", "AM", "fail", "PM"),
                            stat_fn=acc_fun,
                            stat_cols=c("S","R")
)

acc_df$AM_fail <- factor(
  paste0(acc_df$AM, acc_df$fail),
  levels=c( "manualAC", "autoAC","autoAI"),
  labels=c("manual", "AC", "AI"))

```

```{r}
acc_df %>% filter(S %in% c("cc","nn")) %>% ggplot(aes(AM_fail, postmn_stat)) +
  geom_point(size=3)+ geom_errorbar(aes(ymax = upper, ymin = lower), width=0.5) +
  geom_line(aes(group=1), linetype=2)+
  geom_point(aes(y= stat), pch=21, size=4, colour="black")+
  facet_grid(S~PM)  +ylab("Correct RT") +xlab("Automation") +ylab("Accuracy")+
  theme(text = element_text(size = 22), axis.text=element_text(size=15))+ylim(0.5,1)
```

```{r}
acc_df %>% filter(!S %in% c("cc","nn")) %>% ggplot(aes(AM_fail, postmn_stat)) +
  geom_point(size=3)+ geom_errorbar(aes(ymax = upper, ymin = lower), width=0.5) +
  geom_line(aes(group=1), linetype=2)+
  geom_point(aes(y= stat), pch=21, size=4, colour="black")+
  facet_grid(.~S)  +ylab("Correct RT") +xlab("Automation") +ylab("Accuracy")+
  theme(text = element_text(size = 22), axis.text=element_text(size=15)) +ylim(0.5,1)
```

```{r}
corRT_fn_list <- c(function(S, R, rt){quantile(rt[toupper(substr(S,1,1))==R], probs=0.1, na.rm=T)},
                   function(S, R, rt){quantile(rt[toupper(substr(S,1,1))==R], probs=0.5, na.rm=T)},
                   function(S, R, rt){quantile(rt[toupper(substr(S,1,1))==R], probs=0.9, na.rm=T)}
)

cols <- c("S", "R", "rt")
stat_col_list <- list(cols,cols,cols)
stat_fn_names <- c("0.1", "0.5", "0.9")

cRT_df <- stack_stat_postfits(
  dat, pp, facs=c("S", "AM", "fail", "PM"), corRT_fn_list, stat_col_list, stat_fn_names
) 

cRT_df$AM_fail <- factor(
  paste0(cRT_df$AM, cRT_df$fail),
  levels=c( "manualAC", "autoAC","autoAI"),
  labels=c("manual", "AC", "AI"))
```

```{r}
cRT_df %>% filter(S %in% c("cc","nn")) %>%  ggplot(aes(AM_fail, postmn_stat)) +
  geom_point(size=3)+ geom_errorbar(aes(ymax = upper, ymin = lower), width=0.5) +
  geom_line(aes(group=nam), linetype=2)+
  geom_point(aes(y= stat), pch=21, size=4, colour="black")+
  facet_grid(S~PM)  +ylab("Correct RT")  +xlab("Automation")+
  theme(text = element_text(size = 22), axis.text=element_text(size=15))

cRT_df %>% filter(!S %in% c("cc","nn")) %>%  ggplot(aes(AM_fail, postmn_stat)) +
  geom_point(size=3)+ geom_errorbar(aes(ymax = upper, ymin = lower), width=0.5) +
  geom_line(aes(group=nam), linetype=2)+
  geom_point(aes(y= stat), pch=21, size=4, colour="black")+
  facet_grid(.~S)  +ylab("Correct RT")  +xlab("Automation")+
  theme(text = element_text(size = 22), axis.text=element_text(size=15))
```

```{r}
errRT_fn_list <- c(function(S, R, rt){quantile(rt[toupper(substr(S,1,1))!=R],
                                               probs=0.1)},
                   function(S, R, rt){quantile(rt[toupper(substr(S,1,1))!=R], 
                                               probs=0.5)},
                   function(S, R, rt){quantile(rt[toupper(substr(S,1,1))!=R], 
                                               probs=0.9)}
)

eRT_df <- stack_stat_postfits(
  dat, pp, facs=c("S", "AM", "fail", "PM"), errRT_fn_list, stat_col_list, stat_fn_names
)

eRT_df$AM_fail <- factor(
  paste0(eRT_df$AM, eRT_df$fail),
  levels=c( "manualAC", "autoAC","autoAI"),
  labels=c("manual", "AC", "AI"))


eRT_df %>% filter(S %in% c("cc","nn")) %>% ggplot(aes(AM_fail, postmn_stat)) + 
  geom_point(size=3)+ geom_errorbar(aes(ymax = upper, ymin = lower), width=0.5) +
  geom_line(aes(group=nam), linetype=2)+
  geom_point(aes(y= stat), pch=21, size=4, colour="black")+
  facet_grid(S~PM)  +ylab("Error RT") +
  theme(text = element_text(size = 22), axis.text=element_text(size=15))
```

#### **Model Comparison**

Similar model-comparison results to in text. Interestingly, a model with no automation-excitation
specifically in PM conditions is preferred above the most flexible.

```{r}
fit_advanced_noauto <- 
  get(load("samples/sampler_advanced_bound_noauto.RData"))
fit_advanced_noPMinh <- 
  get(load("samples/sampler_advanced_bound_noPMinh.RData"))
fit_advanced_noproactive <- 
  get(load("samples/sampler_advanced_bound_noproactive.RData"))
fit_advanced_noautoexPM <- 
  get(load("samples/sampler_advanced_bound_noautoexPM.RData"))
```

All submodels converged in terms of Rhat <1.1 all parameters.
```{r eval=FALSE}
check(fit_advanced_noauto)
check(fit_advanced_noPMinh)
check(fit_advanced_noproactive)
```

```{r}
out <- compare(list(
  top = fit_advanced,
  noPMinh = fit_advanced_noPMinh,
  noproactive = fit_advanced_noproactive,
  noauto = fit_advanced_noauto,
  noautoex_PM = fit_advanced_noautoexPM
), BayesFactor = FALSE)

out$DIC <- out$DIC - out$DIC[5]
out$BPIC <- out$BPIC - out$BPIC[5]
round(out[,1:4],2)
```

#### **Model Parameter Summaries**

Similar results to in text (except the automation-excitation can't be negative at all
due to the bound)/
```{r}
params <- data.frame(
  get_pars(fit_advanced, selection="mu", merge_chains = T)$mu[[1]]
)
```

```{r}
param_summaries <- params %>% pivot_longer(cols=colnames(params)) %>% 
  group_by(name) %>% 
  summarize(mn=mean(exp(value)), 
            p025= quantile(exp(value), probs=0.025),
            p975=quantile(exp(value), probs=0.975)
  )

exinh <- param_summaries  %>% filter(grepl(".*[ex|inh]$", name))

exinh$name <- c("PM inhibition, auto condition", 
                "PM inhibition, manual condition",
                "Automation excitation, PM condition",
                "Automation inhibition, PM condition",
                "Automation excitation, control condition",
                "Automation inhibition, control condition"
)

exinh  %>% 
  ggplot(aes(x=name, y=mn, ymin=p025, ymax=p975))+
  ylab("exp(Effect)")+ xlab("") +
  geom_pointrange()+ 
  geom_hline(yintercept=0, lty=2) +  # add a dotted line at x=1 after flip
  coord_flip() +
  theme(text = element_text(size = 18))


```

Less Bayesian p-values and Savage-Dickey tests are included here. The simple 
effects with theory-informed bounds can never be 0 or below and therefore there 
is no point testing for that.

Also, tests now involve exponentiating out the pre-transform .
```{r fig.show = "hide"}

set.seed(1234)

PM_control_AutoInh_p <- attr(
  credible(x = fit_advanced, x_fun = function(d) exp(d["v_SlRPMAUTOinh"])
         - exp(d["v_SlRcontrolAUTOinh"])),
  "less"
)
PM_control_AutoInh_SD <- hypothesis(
  fit_advanced, fun=function(d) exp(d["v_SlRPMAUTOinh"])
           - exp(d["v_SlRcontrolAUTOinh"]), do_plot = F)

PM_control_AutoInh_SD

PM_control_AutoEx_p <- attr(
  credible(x = fit_advanced, x_fun = function(d) exp(d["v_SlRcontrolAUTOex"]) 
         - exp(d["v_SlRPMAUTOex"])), "less"
)

PM_control_AutoEx_SD <- hypothesis(fit_advanced, fun= function(d) 
  exp(d["v_SlRcontrolAUTOex"]) - exp(d["v_SlRPMAUTOex"]), do_plot = F)

PM_control_AutoEx_SD 
```


```{r}
proactive <- param_summaries %>% filter(grepl(".*pa.$", name))

proactive$name <- c("Conflict Accumulator, Automation", 
                    "Non-conflict Accumulator, Automation", 
                    "Conflict Accumulator, Manual", 
                    "Non-conflict Accumulator, Manual" 
)

proactive  %>% 
  ggplot(aes(x=name, y=mn, ymin=p025, ymax=p975))+
  ylab("Proactive Control Effect")+ xlab("") +
  geom_pointrange()+ 
  geom_hline(yintercept=0, lty=2) +  # add a dotted line at x=1 after flip
  coord_flip() 

```

```{r fig.show = "hide"}
set.seed(1234)

ProAutoC_p <- attr(credible(x = fit_advanced, x_name= "B_PMlRautopaC"),
                 "less")
ProAutoC_SD <- hypothesis(fit_advanced, parameter="B_PMlRautopaC", do_plot = F)
ProAutoC_SD 

ProAutoN_p <- attr(credible(x = fit_advanced, x_name= "B_PMlRautopaN"),
                 "less")
ProAutoN_SD <- hypothesis(fit_advanced, parameter="B_PMlRautopaN", do_plot = F)
ProAutoN_SD


ProManC_p <- attr(credible(x = fit_advanced, x_name= "B_PMlRmanualpaC"),
                 "less")
ProManC_SD <- hypothesis(fit_advanced, parameter="B_PMlRmanualpaC", do_plot = F)
ProManC_SD

ProManN_p <- attr(credible(x = fit_advanced, x_name= "B_PMlRmanualpaN"),
                 "less")
ProManN_SD <- hypothesis(fit_advanced, parameter="B_PMlRmanualpaN", do_plot = F)
ProManN_SD


ProManAutoC_p <-  attr(credible(x = fit_advanced, x_fun = function(d) d["B_PMlRmanualpaC"] - d["B_PMlRautopaC"]
         ), "less"
)

ProManAutoC_SD <-  hypothesis(fit_advanced, fun = function(d) d["B_PMlRmanualpaC"] - 
             d["B_PMlRautopaC"], do_plot = F
)
ProManAutoC_SD


ProManAutoN_p <-  attr(credible(x = fit_advanced, x_fun = function(d) d["B_PMlRmanualpaN"] - d["B_PMlRautopaN"]
         ), "less"
)

ProManAutoN_SD <-  hypothesis(fit_advanced, fun = function(d) d["B_PMlRmanualpaN"] - 
             d["B_PMlRautopaN"], do_plot = F
)
ProManAutoN_SD

additive_ProManAutoC_p <-  attr(credible(x = fit_advanced, x_fun = function(d) (
 
  (exp(d["B_PMlRmanualC"] +d["B_PMlRmanualpaC"])- exp(d["B_PMlRmanualC"]))  
  - (exp(d["B_PMlRautoC"] +d["B_PMlRautopaC"]) -exp(d["B_PMlRautoC"]))
  )
), "less"
)

additive_ProManAutoC_SD <-  hypothesis(fit_advanced, fun = function(d) (
 
  (exp(d["B_PMlRmanualC"] +d["B_PMlRmanualpaC"])- exp(d["B_PMlRmanualC"]))  
  - (exp(d["B_PMlRautoC"] +d["B_PMlRautopaC"]) -exp(d["B_PMlRautoC"]))
  ), do_plot = F
)
additive_ProManAutoC_SD
  
additive_ProManAutoN_p <-  attr(credible(x = fit_advanced, x_fun = function(d) (
 
  (exp(d["B_PMlRmanualN"] +d["B_PMlRmanualpaN"])- exp(d["B_PMlRmanualN"]))  
  - (exp(d["B_PMlRautoN"] +d["B_PMlRautopaN"]) -exp(d["B_PMlRautoN"]))
  )
), "less"
)

additive_ProManAutoN_SD <-  hypothesis(fit_advanced, fun = function(d) (
 
  (exp(d["B_PMlRmanualN"] +d["B_PMlRmanualpaN"])- exp(d["B_PMlRmanualN"]))  
  - (exp(d["B_PMlRautoN"] +d["B_PMlRautopaN"]) -exp(d["B_PMlRautoN"]))
  ), do_plot = F
)
additive_ProManAutoN_SD 
  

```


